---
title: "RQ2 Data Preparation & Descriptives"
subtitle: "AI Surrogate Sample — Import, Cleaning, Exclusions, Reliability"
author: "Caroline Müller"
date: "2026-05-16"
format:
  html:
    code-fold: true
    code-summary: "Show code"
    code-tools: true
    toc: true
    self-contained: true
    fig-cap-location: top
editor:
  markdown:
    wrap: 72
knitr:
  opts_chunk:
    dpi: 300
---

# 1 Overview

This file covers everything upstream of the inferential RQ2 analyses:

- Raw CSV import and audit
- Type conversion, column renaming, and structural integrity gates
- Sequential exclusion rules (age, completeness, truncation)
- Demographic, textual, and questionnaire descriptives
- LLM score descriptives by condition
- Questionnaire reliability (Cronbach alpha / Omega)
- LLM scoring reliability (ICC)
- CFA on questionnaire items (1-factor vs 3-factor)

**Outputs written to `10_data/80_analyses_exports/`:**

| File | Used by |
|------|---------|
| `d_rq2_clean.rds` | 20_RQ2_analyses |
| `d_rq3_ai_conditions.rds` | 20_RQ2_analyses, 50_additional_analyses |
| `d_rq3_ai.rds` | 30_RQ3_data |
| `tbl_icc_raw.rds` | 30_RQ3_data (display) |
| `tbl_icc2k_raw.rds` | reference |
| `tbl_word_count_by_condition.rds` | 30_RQ3_data |

---

# 2 Setup

## 2.1 Environment

```{r}
#| label: setup-environment
#| include: false
#| warning: false

graphics.off()

suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(tidyr)
  library(flextable)
  library(rempsyc)
  library(scales)
  library(psych)
  library(performance)
  library(officer)
  library(paletteer)
  library(janitor)
  library(systemfonts)
  library(ragg)
  library(lavaan)
  library(stringr)
  library(purrr)
})

knitr::opts_chunk$set(dev = "ragg_png")
if (.Platform$OS.type == "windows") windowsFonts(Calibri = windowsFont("Calibri"))

set.seed(2026)
message("Libraries loaded.")
```

## 2.2 Shared Constants

```{r}
#| label: setup-common
#| include: false

source("00_common.R")
```

## 2.3 File-Level Constants

```{r}
#| label: setup-constants
#| include: false

data_path       <- "../10_data/70_merge_outputs_20260412-1426.csv"
EXPORT_DIR      <- "../10_data/80_analyses_exports"
MAX_LOSS_THRESHOLD <- 0.10

edu_levels <- c("High School", "Some College", "Associate's",
                "Bachelor's", "Master's", "Doctorate", "Other")
```

---

# 3 Data Import

## 3.1 Import CSV

```{r}
#| label: import-raw
#| message: false

d_raw <- read.csv(data_path)
message("Loaded: ", basename(data_path),
        " | Modified: ", format(file.mtime(data_path), "%Y-%m-%d %H:%M"))
```

## 3.2 Import Audit

```{r}
#| label: tbl-import-audit
#| echo: false
#| tbl-cap: "AI Data Import Audit"

tbl_import_audit <- tibble::tribble(
  ~Metric,                   ~Value,
  "Source file",             basename(data_path),
  "Total rows",              as.character(nrow(d_raw)),
  "Distinct AI surrogates",  as.character(dplyr::n_distinct(d_raw$uuid)),
  "Columns",                 as.character(ncol(d_raw))
)

tbl_import_audit |> flextable() |> style_apa7()
```

---

# 4 Data Cleaning

## 4.1 Type Conversion and Renaming

```{r}
#| label: clean-rename
#| include: false

d <- d_raw |>
  rename(
    text_count   = n_texts,
    actual_words = word_count,
    uuid             = uuid,
    age              = age,
    gender           = gender,
    education        = education,
    occupation       = occupation,
    country          = country,
    race             = race,
    nationality      = nationality,
    income_category  = income_category,
    target_words     = target_words,
    repeat_id        = repeat_id,
    q_aut_1 = q_aut_1, q_aut_2 = q_aut_2, q_aut_3 = q_aut_3, q_aut_4 = q_aut_4,
    q_com_1 = q_com_1, q_com_2 = q_com_2, q_com_3 = q_com_3, q_com_4 = q_com_4,
    q_rel_1 = q_rel_1, q_rel_2 = q_rel_2, q_rel_3 = q_rel_3, q_rel_4 = q_rel_4,
    llm_aut = llm_aut, llm_com = llm_com, llm_rel = llm_rel
  ) |>
  mutate(
    uuid            = as.character(uuid),
    age             = as.character(age),
    gender          = as.character(gender),
    education       = as.character(education),
    occupation      = as.character(occupation),
    country         = as.character(country),
    race            = as.character(race),
    nationality     = as.character(nationality),
    income_category = as.character(income_category),
    actual_words    = as.integer(actual_words),
    target_words    = as.integer(target_words),
    text_count      = as.integer(text_count),
    repeat_id       = as.integer(repeat_id),
    llm_aut         = as.integer(llm_aut),
    llm_com         = as.integer(llm_com),
    llm_rel         = as.integer(llm_rel),
    across(starts_with("q_"), as.integer)
  )

message("Type conversion complete.")
```

## 4.2 Structural Integrity Gates

```{r}
#| label: clean-integrity
#| include: false

required_vars <- c("uuid", "text_count", "repeat_id", "actual_words",
                   "llm_aut", "llm_com", "llm_rel")
missing_vars  <- setdiff(required_vars, names(d))
if (length(missing_vars) > 0) {
  stop("CRITICAL: Missing columns: ", paste(missing_vars, collapse = ", "))
}

# Content integrity: flag conditions with > MAX_LOSS_THRESHOLD missing
content_check <- d |>
  group_by(condition_raw = paste0(target_words, "w-", text_count)) |>
  summarise(
    planned    = n(),
    valid_text = sum(!is.na(actual_words)),
    .groups    = "drop"
  )

high_loss <- content_check |>
  filter((planned - valid_text) / planned > MAX_LOSS_THRESHOLD)

if (nrow(high_loss) > 0) {
  stop("CRITICAL DATA LOSS in conditions: ",
       paste(high_loss$condition_raw, collapse = ", "))
}

message("Structural integrity gates passed.")
```

## 4.3 Add Derived Columns

Condition factors, questionnaire composites, and age numeric:

```{r}
#| label: clean-add-columns
#| include: false

# Condition factor
d <- d |>
  mutate(
    condition  = factor(
      paste0(WORD_LENGTH_LABELS[as.character(target_words)], "-", text_count),
      levels = CONDITIONS
    ),
    text_count = factor(text_count, levels = N_TEXTS_VARIANTS)
  )

# Questionnaire composites — exactly 4 items each (AI sample)
d <- d |>
  mutate(
    q_aut_mean = rowMeans(across(c(q_aut_1, q_aut_2, q_aut_3, q_aut_4)), na.rm = TRUE),
    q_com_mean = rowMeans(across(c(q_com_1, q_com_2, q_com_3, q_com_4)), na.rm = TRUE),
    q_rel_mean = rowMeans(across(c(q_rel_1, q_rel_2, q_rel_3, q_rel_4)), na.rm = TRUE)
  )

# Age: parse "20-29" style strings into age_low, age_high, age_mid, age_bracket
process_age_string <- function(x) {
  x_char <- trimws(as.character(x))
  lower  <- NA_real_; upper <- NA_real_
  if (grepl("-", x_char)) {
    parts <- as.numeric(strsplit(x_char, "\\s*-\\s*")[[1]])
    lower <- parts[1]; upper <- parts[2]
  } else if (x_char == "70 or older") {
    lower <- 70; upper <- 75
  } else if (grepl("^[0-9]+$", x_char)) {
    lower <- as.numeric(x_char); upper <- as.numeric(x_char)
  }
  data.frame(age_low = lower, age_high = upper, age_mid = (lower + upper) / 2)
}

age_raw <- d |> distinct(age) |> filter(!is.na(age)) |>
  mutate(age = as.character(age)) |> pull(age)

age_map <- purrr::map(age_raw, function(v) {
  res          <- process_age_string(v)
  d_start      <- floor(res$age_low / 10) * 10
  res$age_bracket <- if (!is.na(res$age_low) && res$age_low >= 70) "70 or older" else
                        paste0(d_start, "-", d_start + 9)
  res$decade_start <- d_start
  data.frame(age = v, res, stringsAsFactors = FALSE)
}) |> purrr::list_rbind()

d <- d |>
  mutate(age = as.character(age)) |>
  dplyr::select(-any_of(c("age_low", "age_high", "age_mid", "age_bracket",
                           "decade_start"))) |>
  left_join(age_map, by = "age")

rm(age_raw)
message("Derived columns added.")
```

---

# 5 Exclusion Rules

```{r}
#| label: exclusion-init
#| include: false

n_step0    <- n_distinct(d$uuid); rows_step0 <- nrow(d)
message("Step 0 (import): ", n_step0, " AI surrogates, ", rows_step0, " rows.")
```

## 5.1 Rule 1 — Age (20–69)

```{r}
#| label: exclusion-rule1-age
#| include: false

d <- d |> filter(age_low >= 20, age_low < 70)
n_step1    <- n_distinct(d$uuid); rows_step1 <- nrow(d)
message("Rule 1 (age): N = ", n_step1, " remaining.")
```

## 5.2 Rule 2 — Row Completeness (27 rows per persona)

```{r}
#| label: exclusion-rule2-completeness
#| include: false

incomplete_uuids <- d |> count(uuid) |>
  filter(n != ROWS_PER_PERSONA) |> pull(uuid)
n_removed_incomplete <- length(incomplete_uuids)

# Rule 2 filter is suspended if all personas are complete; kept commented for
# production confirmation — uncomment to enforce:
# d <- d |> filter(!(uuid %in% incomplete_uuids))

n_step2    <- n_distinct(d$uuid); rows_step2 <- nrow(d)
if (n_removed_incomplete > 0) {
  message("Rule 2 (completeness): ", n_removed_incomplete,
          " incomplete persona(s) found — filter suspended.")
} else {
  message("Rule 2 (completeness): All personas have ", ROWS_PER_PERSONA, " rows.")
}
```

## 5.3 Rule 3 — Text Truncation (< 75% of target)

```{r}
#| label: exclusion-rule3-truncation
#| include: false

uuids_failing <- d |>
  filter(actual_words < (target_words * TEXT_TRUNC_THRESHOLD)) |>
  distinct(uuid) |> pull(uuid)

n_personas_removed_trunc <- length(uuids_failing)
n_rows_removed_trunc     <- sum(d$uuid %in% uuids_failing)
d <- d |> filter(!uuid %in% uuids_failing)

n_step3    <- n_distinct(d$uuid); rows_step3 <- nrow(d)
message("Rule 3 (truncation): Removed ", n_personas_removed_trunc,
        " UUID(s) (", n_rows_removed_trunc, " rows). Final N = ", n_step3, ".")
```

## 5.4 Attrition Table

```{r}
#| label: tbl-attrition
#| echo: false
#| tbl-cap: "Sequential Exclusion Rules Applied to AI Surrogate Sample"

tbl_attrition <- tibble::tibble(
  Step = 0:3,
  Rule = c("Raw import", "Age (20–69)", "Completeness (27 rows)", "Truncation (75% threshold)"),
  `Removed (Pers.)` = c("—", n_step0 - n_step1, n_step1 - n_step2, n_step2 - n_step3),
  `Removed (Rows)`  = c("—", rows_step0 - rows_step1, rows_step1 - rows_step2, rows_step2 - rows_step3),
  `AI Surrogates (n)` = c(n_step0, n_step1, n_step2, n_step3),
  `Rows (n)` = c(rows_step0, rows_step1, rows_step2, rows_step3)
)

tbl_attrition |>
  flextable() |>
  align(j = 1:6, align = "center", part = "all") |>
  style_apa7() |>
  bold(i = ~ as.numeric(`Removed (Pers.)`) > 0, j = 3:4) |>
  add_apa_note("Steps 1–3 are surrogate-level exclusions. Rule 2 (completeness) filter suspended pending production confirmation.")
```

## 5.5 Final Sample

```{r}
#| label: final-sample-summary
#| include: false

d_persona <- d |> distinct(uuid, .keep_all = TRUE)

# Verify no truncated rows remain
n_trunc_rem <- sum(d$actual_words < (d$target_words * TEXT_TRUNC_THRESHOLD), na.rm = TRUE)
if (n_trunc_rem > 0) warning("Rule 3 Alert: ", n_trunc_rem, " truncated rows remain.")

message("Final analytic sample: ", nrow(d_persona), " AI surrogates, ",
        nrow(d), " rows.")
```

```{r}
#| label: setup-d-conditions
#| include: false

# Aggregate 3 scoring repeats to per-persona × condition means
n_surviving     <- n_distinct(d$uuid)
expected_rows   <- n_surviving * length(CONDITIONS)

d_conditions <- d |>
  group_by(uuid, condition) |>
  summarise(
    text_count   = first(text_count),
    target_words = first(target_words),
    actual_words = mean(actual_words, na.rm = TRUE),  # mean across repeats (pool varies per text_count)
    across(c(llm_aut, llm_com, llm_rel), \(x) mean(x, na.rm = TRUE), .names = "{.col}_m"),
    n_valid      = sum(!is.na(llm_aut)),
    .groups      = "drop"
  )

stopifnot("d_conditions row count mismatch" = nrow(d_conditions) == expected_rows)
message("d_conditions created: ", nrow(d_conditions), " rows (", n_surviving,
        " personas × ", length(CONDITIONS), " conditions).")
```

---

# 6 Descriptives: Demographics

## 6.1 Age

```{r}
#| label: tbl-age-dist
#| echo: false
#| tbl-cap: "Distribution of AI Surrogates by Age Bracket"

persona_stats <- d_persona |>
  summarise(
    Mean_Age = mean(age_mid, na.rm = TRUE),
    SD_Age   = sd(age_mid,   na.rm = TRUE),
    Min_Age  = min(age_low,  na.rm = TRUE),
    Max_Age  = max(age_high, na.rm = TRUE)
  )

tbl_age_dist <- d_persona |>
  count(age_bracket, decade_start) |>
  arrange(decade_start) |>
  dplyr::select(`Age Bracket` = age_bracket, n) |>
  mutate(pct = scales::percent(n / sum(n), accuracy = 0.1)) |>
  janitor::adorn_totals("row") |>
  as_tibble() |>
  mutate(pct = if_else(`Age Bracket` == "Total",
                       scales::percent(1, accuracy = 0.1), pct))

tbl_age_dist |>
  flextable() |>
  style_apa7() |>
  bold(i = ~ `Age Bracket` == "Total", j = c("n", "pct")) |>
  set_header_labels(pct = "%") |>
  add_apa_note(sprintf(
    "Mean age (midpoint estimate) = %.1f years (SD = %.1f; range: %d–%d).",
    persona_stats$Mean_Age, persona_stats$SD_Age,
    as.integer(persona_stats$Min_Age), as.integer(persona_stats$Max_Age)
  ))
```

## 6.2 Gender

```{r}
#| label: tbl-gender-dist
#| echo: false
#| tbl-cap: "Gender Distribution of AI Surrogate Sample"

d_persona |>
  count(Gender = gender) |>
  janitor::adorn_totals("row") |>
  mutate(pct = scales::percent(n / sum(n[Gender != "Total"]), accuracy = 0.1)) |>
  flextable() |>
  style_apa7() |>
  bold(i = ~ Gender == "Total", j = c("n", "pct")) |>
  set_header_labels(pct = "%")
```

## 6.3 Education

```{r}
#| label: setup-education-clean
#| include: false

d_persona <- d_persona |>
  mutate(education_clean = case_match(
    trimws(tolower(education)),
    c("high school", "highschool")   ~ "High School",
    c("some college", "somecollege") ~ "Some College",
    c("associate's", "associates")   ~ "Associate's",
    c("bachelor's", "bachelors")     ~ "Bachelor's",
    c("master's", "masters")         ~ "Master's",
    c("doctorate", "phd")            ~ "Doctorate",
    .default = "Other"
  )) |>
  mutate(education_clean = factor(education_clean, levels = edu_levels))

if (any(is.na(d_persona$education_clean))) {
  warning("Unmapped strings detected in education field.")
}
```

```{r}
#| label: tbl-education-dist
#| echo: false
#| tbl-cap: "Education Level Distribution of AI Surrogate Sample"

d_persona |>
  count(Education = education_clean) |>
  janitor::adorn_totals("row") |>
  mutate(pct = scales::percent(n / sum(n[Education != "Total"]), accuracy = 0.1)) |>
  flextable() |>
  style_apa7() |>
  bold(i = ~ Education == "Total", j = c("n", "pct")) |>
  set_header_labels(Education = "Education Level", pct = "%")
```

---

# 7 Descriptives: Textual Properties

## 7.1 Word Counts by Condition

```{r}
#| label: tbl-wordcount-by-condition
#| echo: false
#| tbl-cap: "Descriptive Statistics for Actual Word Counts by Experimental Condition"

tbl_word_count_by_condition <- d |>
  group_by(Condition = condition, `Target (w)` = target_words,
           `Texts (n)` = text_count) |>
  summarise(
    M   = round(mean(actual_words, na.rm = TRUE), 1),
    SD  = round(sd(actual_words,   na.rm = TRUE), 1),
    Min = min(actual_words,  na.rm = TRUE),
    Max = max(actual_words,  na.rm = TRUE),
    n   = n(),
    .groups = "drop"
  )

tbl_word_count_by_condition |>
  flextable() |>
  style_apa7() |>
  add_apa_note("Word count is per text (single text per row). N reflects all rows per condition across repeats.")
```

## 7.2 Truncation Compliance

```{r}
#| label: tbl-truncation-audit
#| echo: false
#| tbl-cap: "Text Length Compliance by Target Word Count (75% Threshold)"

d |>
  group_by(`Target (w)` = target_words) |>
  summarise(
    `Planned (N)` = n(),
    `Failed (n)`  = sum(actual_words < (target_words * TEXT_TRUNC_THRESHOLD), na.rm = TRUE),
    `Compliant (n)` = `Planned (N)` - `Failed (n)`,
    `Compliance (%)` = scales::percent(`Compliant (n)` / `Planned (N)`, accuracy = 0.1),
    .groups = "drop"
  ) |>
  flextable() |>
  style_apa7() |>
  add_apa_note(paste0("Compliance = actual_words ≥ ", TEXT_TRUNC_THRESHOLD * 100,
                      "% of target. N aggregated across text-count variants and repeats."))
```

---

# 8 Descriptives: Questionnaire Scores

## 8.1 Item-Level Statistics

```{r}
#| label: tbl-q-items
#| echo: false
#| tbl-cap: "Questionnaire Item-Level Descriptives (AI Surrogate Sample)"

q_item_stats <- d_persona |>
  dplyr::select(q_aut_1:q_rel_4) |>
  summarise(across(everything(), list(
    M  = \(x) round(mean(x,   na.rm = TRUE), 2),
    SD = \(x) round(sd(x,     na.rm = TRUE), 2),
    Md = \(x) round(median(x, na.rm = TRUE), 2),
    n  = \(x) sum(!is.na(x))
  ))) |>
  tidyr::pivot_longer(everything(),
    names_to  = c("item", ".value"),
    names_sep = "_(?=[^_]+$)"
  ) |>
  mutate(
    Subscale = case_when(
      startsWith(item, "q_aut") ~ TABLE_LABEL_AUT,
      startsWith(item, "q_com") ~ TABLE_LABEL_COM,
      startsWith(item, "q_rel") ~ TABLE_LABEL_REL
    ),
    Item = sub("^q_", "", item)
  ) |>
  dplyr::select(Subscale, Item, M, SD, Md, n)

q_item_stats |>
  flextable() |>
  style_apa7() |>
  add_apa_note(paste0("N = ", nrow(d_persona), " AI surrogates. Items scored 1–5."))
```

## 8.2 Composite Means by Need

```{r}
#| label: tbl-q-composites
#| echo: false
#| tbl-cap: "Questionnaire Composite Score Descriptives by Need Dimension"

q_comp_stats <- d_persona |>
  dplyr::select(q_aut_mean, q_com_mean, q_rel_mean) |>
  tidyr::pivot_longer(everything(), names_to = "subscale", values_to = "score") |>
  mutate(subscale = case_match(subscale,
    "q_aut_mean" ~ TABLE_LABEL_AUT,
    "q_com_mean" ~ TABLE_LABEL_COM,
    "q_rel_mean" ~ TABLE_LABEL_REL
  )) |>
  group_by(subscale) |>
  summarise(
    M   = round(mean(score,   na.rm = TRUE), 2),
    SD  = round(sd(score,     na.rm = TRUE), 2),
    Md  = round(median(score, na.rm = TRUE), 2),
    Min = round(min(score,    na.rm = TRUE), 2),
    Max = round(max(score,    na.rm = TRUE), 2),
    n   = sum(!is.na(score)),
    .groups = "drop"
  )

q_comp_stats |>
  rename(!!LABEL_NEEDS := subscale) |>
  flextable() |>
  style_apa7() |>
  add_apa_note(paste0("N = ", nrow(d_persona), " AI surrogates. Composite = mean of 4 items per need."))
```

## 8.3 Norms Comparison (Olafsen et al., 2021)

```{r}
#| label: tbl-norm-comparison
#| echo: false
#| tbl-cap: "Comparison of AI Surrogate Questionnaire Scores with Human Norms (Olafsen et al., 2021)"

# Olafsen et al. (2021), Sample 4 (N = 393), rescaled from 1–7 to 1–5
olafsen_norms <- tribble(
  ~subscale,    ~norm_m, ~norm_sd,
  "autonomy",     3.52,     0.59,
  "competence",   3.86,     0.56,
  "relatedness",  3.76,     0.63
)

tbl_norm_comp <- olafsen_norms |>
  left_join(q_comp_stats |>
              mutate(subscale = tolower(case_match(subscale,
                TABLE_LABEL_AUT ~ "autonomy",
                TABLE_LABEL_COM ~ "competence",
                TABLE_LABEL_REL ~ "relatedness"
              ))), by = "subscale") |>
  mutate(
    `Delta M`  = round(M - norm_m,  2),
    `Delta SD` = round(SD - norm_sd, 2),
    subscale   = tools::toTitleCase(subscale)
  ) |>
  dplyr::select(
    `Need Dimension` = subscale,
    `Norm M`         = norm_m,
    `Norm SD`        = norm_sd,
    `Surrogate M`    = M,
    `Surrogate SD`   = SD,
    `Delta M`, `Delta SD`
  )

tbl_norm_comp |>
  flextable() |>
  style_apa7() |>
  add_apa_note(paste0(
    "Norms from Olafsen et al. (2021), Sample 4 (N = 393), rescaled from 1–7 to 1–5 scale. ",
    "AI Surrogate N = ", nrow(d_persona), "."
  ))
```

---

# 9 Descriptives: LLM Scores

## 9.1 LLM Score Distributions by Condition

```{r}
#| label: tbl-llm-by-condition
#| echo: false
#| tbl-cap: "LLM Score Descriptives by Condition and Need Dimension (Surrogate-Level Means)"

d_conditions |>
  tidyr::pivot_longer(c(llm_aut_m, llm_com_m, llm_rel_m),
                      names_to = "need", values_to = "score") |>
  mutate(need = case_match(need,
    "llm_aut_m" ~ TABLE_LABEL_AUT,
    "llm_com_m" ~ TABLE_LABEL_COM,
    "llm_rel_m" ~ TABLE_LABEL_REL
  )) |>
  group_by(Condition = condition, Need = need) |>
  summarise(
    M   = round(mean(score, na.rm = TRUE), 2),
    SD  = round(sd(score,   na.rm = TRUE), 2),
    Min = round(min(score,  na.rm = TRUE), 2),
    Max = round(max(score,  na.rm = TRUE), 2),
    n   = n(),
    .groups = "drop"
  ) |>
  flextable() |>
  style_apa7() |>
  add_apa_note(paste0("N per condition = ", n_surviving, " AI surrogates. ",
                      "Scores = mean of 3 scoring repeats per persona."))
```

## 9.2 LLM Score Distribution — Boxplots by Condition

```{r}
#| label: fig-llm-condition-boxplot
#| echo: false
#| fig-cap: "LLM Score Distributions by Condition and Need Dimension (Surrogate-Level Means). Each box represents one experimental condition (target word length × number of texts)."
#| fig-width: 9
#| fig-height: 5
#| fig-dpi: 300

d_conditions |>
  tidyr::pivot_longer(c(llm_aut_m, llm_com_m, llm_rel_m),
                      names_to = "need", values_to = "score") |>
  mutate(need = case_match(need,
    "llm_aut_m" ~ TABLE_LABEL_AUT,
    "llm_com_m" ~ TABLE_LABEL_COM,
    "llm_rel_m" ~ TABLE_LABEL_REL
  )) |>
  ggplot(aes(x = condition, y = score, fill = need)) +
  geom_boxplot(alpha = 0.7, outlier.size = 0.8) +
  scale_fill_manual(values = COL_AI_LLM, name = "Need") +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(x = "Condition", y = "LLM Score (1–5)") +
  theme_thesis()
```

## 9.3 LLM vs Questionnaire Score Overlay

```{r}
#| label: fig-llm-q-overlay
#| echo: false
#| fig-cap: "LLM vs Questionnaire Score Distributions by Need Dimension. Blue = LLM scores; gray = questionnaire scores."
#| fig-width: 8
#| fig-height: 4
#| fig-dpi: 300

overlay_data <- bind_rows(
  d_conditions |>
    dplyr::select(uuid, condition, llm_aut_m, llm_com_m, llm_rel_m) |>
    tidyr::pivot_longer(c(llm_aut_m, llm_com_m, llm_rel_m),
                        names_to = "need", values_to = "score") |>
    mutate(
      source = "LLM",
      need   = case_match(need,
        "llm_aut_m" ~ TABLE_LABEL_AUT,
        "llm_com_m" ~ TABLE_LABEL_COM,
        "llm_rel_m" ~ TABLE_LABEL_REL
      )
    ),
  d_persona |>
    dplyr::select(uuid, q_aut_mean, q_com_mean, q_rel_mean) |>
    tidyr::pivot_longer(c(q_aut_mean, q_com_mean, q_rel_mean),
                        names_to = "need", values_to = "score") |>
    mutate(
      source = "Questionnaire",
      need   = case_match(need,
        "q_aut_mean" ~ TABLE_LABEL_AUT,
        "q_com_mean" ~ TABLE_LABEL_COM,
        "q_rel_mean" ~ TABLE_LABEL_REL
      )
    )
)

ggplot(overlay_data, aes(x = score, fill = source)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~need, ncol = 3) +
  scale_fill_manual(values = c(LLM = COL_AI_SINGLE, Questionnaire = COL_NEUTRAL_SINGLE),
                    name = "Measure") +
  labs(x = "Score (1–5)", y = "Density") +
  theme_thesis()
```

---

# 10 Reliability

## 10.1 Questionnaire Reliability (Cronbach Alpha / Omega)

```{r}
#| label: tbl-reliability-omega
#| echo: false
#| cache: true
#| cache.extra: !expr file.mtime(data_path)
#| tbl-cap: "Internal Consistency of Questionnaire Subscales (Cronbach's Alpha and McDonald's Omega)"

omega_aut <- psych::omega(d_persona |> dplyr::select(q_aut_1:q_aut_4), plot = FALSE)
omega_com <- psych::omega(d_persona |> dplyr::select(q_com_1:q_com_4), plot = FALSE)
omega_rel <- psych::omega(d_persona |> dplyr::select(q_rel_1:q_rel_4), plot = FALSE)

alpha_aut <- psych::alpha(d_persona |> dplyr::select(q_aut_1:q_aut_4))$total$raw_alpha
alpha_com <- psych::alpha(d_persona |> dplyr::select(q_com_1:q_com_4))$total$raw_alpha
alpha_rel <- psych::alpha(d_persona |> dplyr::select(q_rel_1:q_rel_4))$total$raw_alpha

tbl_reliability <- tibble::tibble(
  Need  = NEEDS,
  Items = 4L,
  Alpha = round(c(alpha_aut, alpha_com, alpha_rel), 2),
  Omega = round(c(omega_aut$omega.tot, omega_com$omega.tot, omega_rel$omega.tot), 2)
)

tbl_reliability |>
  flextable() |>
  style_apa7() |>
  add_apa_note(paste0("N = ", nrow(d_persona), ". Alpha = Cronbach's alpha (raw). ",
                      "Omega = McDonald's omega (total). Items scored 1–5."))
```

## 10.2 LLM Scoring Reliability — ICC(2,1) and ICC(2,3)

ICC(2,1) characterises single-draw reliability; ICC(2,3) characterises the
reliability of the mean across k = 3 scoring repeats, which is the
quantity used in all downstream analyses. In a small number of condition × need
cells the underlying lme4 model hit the parameter boundary (random-effect
variance estimated at zero), which is expected when ICC approaches its theoretical
limits and does not affect the reported values.

```{r}
#| label: tbl-llm-icc-results
#| echo: false
#| warning: false
#| cache: true
#| cache.extra: !expr file.mtime(data_path)
#| tbl-cap: "LLM Scoring Reliability: ICC(2,1) — Two-Way Random Effects, Absolute Agreement, Single Measures"

calc_condition_icc <- function(sub_data, score_var) {
  wide_reps <- sub_data |>
    dplyr::select(uuid, repeat_id, all_of(score_var)) |>
    filter(!is.na(.data[[score_var]])) |>
    tidyr::pivot_wider(names_from = repeat_id, names_prefix = "rep",
                       values_from = all_of(score_var))

  n_expected <- n_distinct(sub_data$uuid)
  if (nrow(wide_reps) != n_expected) {
    warning("ICC pivot: expected ", n_expected, " rows, got ", nrow(wide_reps),
            " for ", score_var)
  }

  icc_out <- psych::ICC(wide_reps |> dplyr::select(starts_with("rep")),
                        missing = TRUE)

  icc2  <- icc_out$results |> filter(type == "ICC2")
  icc2k <- icc_out$results |> filter(type == "ICC2k")

  list(
    icc2_val  = icc2$ICC,   icc2_lo  = icc2$`lower bound`,  icc2_hi  = icc2$`upper bound`,
    icc2k_val = icc2k$ICC, icc2k_lo = icc2k$`lower bound`, icc2k_hi = icc2k$`upper bound`
  )
}

conditions_grid <- d |>
  distinct(condition, target_words, text_count) |>
  arrange(condition)

tbl_icc_raw <- purrr::map_dfr(seq_len(nrow(conditions_grid)), function(i) {
  cond <- conditions_grid$condition[i]
  sub  <- d |> filter(condition == cond)
  purrr::map_dfr(c("llm_aut", "llm_com", "llm_rel"), function(sv) {
    stats <- calc_condition_icc(sub, sv)
    tibble::tibble(
      Condition = as.character(cond), Need = sv,
      ICC    = stats$icc2_val,
      CI_lo  = stats$icc2_lo,
      CI_hi  = stats$icc2_hi
    )
  })
})

tbl_icc_raw |>
  mutate(
    Need   = case_match(Need, "llm_aut" ~ TABLE_LABEL_AUT,
                               "llm_com" ~ TABLE_LABEL_COM,
                               "llm_rel" ~ TABLE_LABEL_REL),
    Result = sprintf("%.2f [%.2f, %.2f]", ICC, CI_lo, CI_hi)
  ) |>
  dplyr::select(Condition, Need, Result) |>
  tidyr::pivot_wider(names_from = Need, values_from = Result) |>
  flextable() |>
  style_apa7() |>
  add_header_row(
    values     = c("", "ICC(2,1) [95% CI]"),
    colwidths  = c(1, 3)
  ) |>
  align(i = 1, align = "center", part = "header") |>
  add_apa_note(paste0(
    "Two-way random effects model, absolute agreement, single measures. ",
    "Computed via psych::ICC with missing = TRUE."
  ))
```

```{r}
#| label: tbl-llm-icc2k-results
#| echo: false
#| warning: false
#| cache: true
#| cache.extra: !expr file.mtime(data_path)
#| tbl-cap: "LLM Scoring Reliability: ICC(2,3) — Two-Way Random Effects, Absolute Agreement, Average of k = 3 Measures"

tbl_icc2k_raw <- purrr::map_dfr(seq_len(nrow(conditions_grid)), function(i) {
  cond <- conditions_grid$condition[i]
  sub  <- d |> filter(condition == cond)
  purrr::map_dfr(c("llm_aut", "llm_com", "llm_rel"), function(sv) {
    stats <- calc_condition_icc(sub, sv)
    tibble::tibble(
      Condition = as.character(cond), Need = sv,
      ICC    = stats$icc2k_val,
      CI_lo  = stats$icc2k_lo,
      CI_hi  = stats$icc2k_hi
    )
  })
})

# Reliability alert (Layer C)
low_reliability <- tbl_icc_raw |> filter(ICC < 0.75)
if (nrow(low_reliability) > 0) {
  message("Reliability Alert: ", nrow(low_reliability),
          " condition × need cell(s) with ICC(2,1) < .75:")
  print(low_reliability)
} else {
  message("Reliability OK: All ICC(2,1) values ≥ .75.")
}

tbl_icc2k_raw |>
  mutate(
    Need   = case_match(Need, "llm_aut" ~ TABLE_LABEL_AUT,
                               "llm_com" ~ TABLE_LABEL_COM,
                               "llm_rel" ~ TABLE_LABEL_REL),
    Result = sprintf("%.2f [%.2f, %.2f]", ICC, CI_lo, CI_hi)
  ) |>
  dplyr::select(Condition, Need, Result) |>
  tidyr::pivot_wider(names_from = Need, values_from = Result) |>
  flextable() |>
  style_apa7() |>
  add_header_row(
    values     = c("", "ICC(2,3) [95% CI]"),
    colwidths  = c(1, 3)
  ) |>
  align(i = 1, align = "center", part = "header") |>
  add_apa_note(paste0(
    "Two-way random effects model, absolute agreement, average of k = 3 measures (Shrout & Fleiss ICC(2,k)). ",
    "The three scoring repeats are treated as exchangeable raters. ",
    "ICC(2,3) reflects reliability of the mean score used in downstream analyses."
  ))
```

## 10.3 CFA on Questionnaire Items (1-Factor vs 3-Factor)

We fit a 1-factor (general valence) and 3-factor (SDT need structure) CFA on the
12 questionnaire items (one row per persona) and compare fit indices. The
chi-square difference test between these models is an approximation: strict
nesting would require constraining the 3-factor inter-factor correlations to 1.0;
the LRT here serves as a descriptive comparison of fit improvement only.

```{r}
#| label: validate-q-cfa
#| include: false
#| message: false

q_model_1f <- '
  General =~ q_aut_1 + q_aut_2 + q_aut_3 + q_aut_4 +
             q_com_1 + q_com_2 + q_com_3 + q_com_4 +
             q_rel_1 + q_rel_2 + q_rel_3 + q_rel_4
'

q_model_3f <- '
  Autonomy    =~ q_aut_1 + q_aut_2 + q_aut_3 + q_aut_4
  Competence  =~ q_com_1 + q_com_2 + q_com_3 + q_com_4
  Relatedness =~ q_rel_1 + q_rel_2 + q_rel_3 + q_rel_4
'

fit_1f <- lavaan::cfa(q_model_1f, data = d_persona)
fit_3f <- lavaan::cfa(q_model_3f, data = d_persona)

idx_names <- c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr")
fi_1f <- lavaan::fitmeasures(fit_1f, idx_names)
fi_3f <- lavaan::fitmeasures(fit_3f, idx_names)

tbl_cfa_fit <- rbind(`1-factor` = round(fi_1f, 3),
                     `3-factor` = round(fi_3f, 3)) |>
  as.data.frame() |>
  tibble::rownames_to_column("Model")

lrt_cfa <- lavaan::lavTestLRT(fit_1f, fit_3f)

cat("LRT chi-square difference (3-factor vs 1-factor, approximation):\n")
print(lrt_cfa)
```

```{r}
#| label: tbl-cfa-fit
#| echo: false
#| tbl-cap: "CFA Model Fit: 1-Factor vs 3-Factor Questionnaire Structure"

tbl_cfa_fit |>
  flextable() |>
  set_header_labels(chisq = "chi-sq", pvalue = "p", cfi = "CFI",
                    rmsea = "RMSEA", srmr = "SRMR") |>
  style_apa7() |>
  add_apa_note(paste0(
    "N = ", nrow(d_persona), " AI surrogates. ",
    "Chi-square difference test between models is an approximation (models are not ",
    "strictly nested without constraining 3-factor inter-correlations to 1.0)."
  ))
```

---

# 11 Exports

```{r}
#| label: export-all
#| include: false

# d_rq2_clean: long form, N_personas x 27 rows (9 conditions x 3 repeats)
# contains raw llm_aut/com/rel per repeat, questionnaire items, condition factor,
# q_aut/com/rel_mean composites, and age_mid/age_bracket derived columns
d_rq2_clean <- d
saveRDS(d_rq2_clean,
        file.path(EXPORT_DIR, "d_rq2_clean.rds"))
message("Exported: d_rq2_clean.rds (", nrow(d_rq2_clean), " rows)")

# d_rq3_ai: single matched condition (short-3) for human merge
RQ3_EXPORT_CONDITION <- "short-3"
stopifnot("RQ3 export condition not found" =
            RQ3_EXPORT_CONDITION %in% d_conditions$condition)

d_rq3_ai <- d_conditions |>
  filter(condition == RQ3_EXPORT_CONDITION) |>
  left_join(d_persona |> dplyr::select(uuid, q_aut_mean, q_com_mean, q_rel_mean),
            by = "uuid") |>
  mutate(id_type = "AI") |>
  rename(id = uuid, llm_aut_mean = llm_aut_m,
         llm_com_mean = llm_com_m, llm_rel_mean = llm_rel_m) |>
  dplyr::select(id, id_type, llm_aut_mean, llm_com_mean, llm_rel_mean,
                q_aut_mean, q_com_mean, q_rel_mean, actual_words,
                condition, text_count, target_words)

saveRDS(d_rq3_ai, file.path(EXPORT_DIR, "d_rq3_ai.rds"))
message("Exported: d_rq3_ai.rds (", nrow(d_rq3_ai), " rows, condition = ",
        RQ3_EXPORT_CONDITION, ")")
rm(d_rq3_ai)

# d_rq3_ai_conditions: all 9 conditions (for RQ2 analyses and additional analyses)
d_rq3_ai_conditions <- d_conditions |>
  left_join(d_persona |> dplyr::select(uuid, q_aut_mean, q_com_mean, q_rel_mean),
            by = "uuid") |>
  mutate(id_type = "AI") |>
  rename(id = uuid, llm_aut_mean = llm_aut_m,
         llm_com_mean = llm_com_m, llm_rel_mean = llm_rel_m) |>
  dplyr::select(id, id_type, llm_aut_mean, llm_com_mean, llm_rel_mean,
                q_aut_mean, q_com_mean, q_rel_mean, actual_words,
                condition, text_count, target_words)

saveRDS(d_rq3_ai_conditions, file.path(EXPORT_DIR, "d_rq3_ai_conditions.rds"))
message("Exported: d_rq3_ai_conditions.rds (", nrow(d_rq3_ai_conditions), " rows, all 9 conditions)")
rm(d_rq3_ai_conditions)

# tbl_icc_raw and tbl_icc2k_raw (computed in cached chunks above)
saveRDS(tbl_icc_raw,   file.path(EXPORT_DIR, "tbl_icc_raw.rds"))
saveRDS(tbl_icc2k_raw, file.path(EXPORT_DIR, "tbl_icc2k_raw.rds"))
message("Exported: tbl_icc_raw.rds, tbl_icc2k_raw.rds")

# tbl_word_count_by_condition: raw tibble (not the flextable)
tbl_word_count_rds <- d |>
  dplyr::group_by(condition, target_words, text_count) |>
  dplyr::summarise(
    mean_words = mean(actual_words, na.rm = TRUE),
    sd_words   = sd(actual_words,   na.rm = TRUE),
    n          = dplyr::n(),
    .groups    = "drop"
  )
saveRDS(tbl_word_count_rds, file.path(EXPORT_DIR, "tbl_word_count_by_condition.rds"))
message("Exported: tbl_word_count_by_condition.rds")

message("All exports complete.")
```

---

# 12 Session Info

```{r}
#| label: session-info
#| include: false

sessionInfo()
```
