---
title: "H1c & H1d: Discriminant Validity and Structural Consistency"
author: "Caroline Müller"
format:
  html:
    code-fold: true
    code-summary: "Show code"
    code-tools: true
    toc: true
    self-contained: true
date: "2026-05-07"
editor:
  markdown:
    wrap: 72
knitr:
  opts_chunk:
    dpi: 300
---

Analyses for H1c (discriminant validity) and H1d (structural consistency
of inter-need patterns). Loads the pre-processed per-condition dataset
exported by `10_Analyses_RQ2.qmd` — all exclusions and repeat-aggregations
are already applied.

Input: `10_data/80_analyses_exports/d_rq3_ai_conditions.rds`

# 1 Setup

```{r}
#| label: setup-libraries
#| include: false

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(purrr)
  library(ggplot2)
  library(systemfonts)
  library(ragg)
  library(flextable)
  library(officer)
  library(cocor)
})

knitr::opts_chunk$set(dev = "ragg_png")
if (.Platform$OS.type == "windows") windowsFonts(Calibri = windowsFont("Calibri"))

# Resolve tidyverse conflicts explicitly
conflicted::conflicts_prefer(dplyr::filter, dplyr::lag, .quiet = TRUE)
```

```{r}
#| label: setup-constants
#| include: false

DATA_PATH <- "../10_data/80_analyses_exports/d_rq3_ai_conditions.rds"

# APA-style flextable helper (mirrors RQ2.qmd)
style_apa7 <- function(ft) {
  ft |>
    flextable::font(fontname = "Times New Roman", part = "all") |>
    flextable::fontsize(size = 12, part = "all") |>
    flextable::border_remove() |>
    flextable::hline_top(part = "header",
                         border = officer::fp_border(width = 1.5)) |>
    flextable::hline_bottom(part = "header",
                            border = officer::fp_border(width = 1)) |>
    flextable::hline_bottom(part = "body",
                            border = officer::fp_border(width = 1.5)) |>
    flextable::align(align = "left", part = "all") |>
    flextable::padding(padding = 3, part = "all")
}
```

# 2 Data

```{r}
#| label: import-h1-data
#| include: false

d_h1 <- readRDS(DATA_PATH)

# Structural gate
stopifnot(
  all(c("q_aut_mean", "q_com_mean", "q_rel_mean",
        "llm_aut_mean", "llm_com_mean", "llm_rel_mean",
        "target_words", "text_count") %in% names(d_h1)),
  nrow(d_h1) > 0
)

# One row per persona × condition
conditions <- d_h1 |>
  distinct(target_words, text_count) |>
  arrange(text_count, target_words)

cat("Loaded:", nrow(d_h1), "rows |",
    n_distinct(d_h1$id), "personas |",
    nrow(conditions), "conditions\n")
```

# 3 H1c: Discriminant Validity

Discriminant validity requires that the LLM score for a given need
dimension correlates more strongly with the *matching* questionnaire
dimension (r_within) than with *non-matching* dimensions (r_cross).

Because all three correlations share the same sample and a common
variable (the LLM score), they are **dependent**. Steiger's (1980)
test for overlapping dependent correlations is used rather than the
standard Fisher r-to-z transform (which assumes independence). This
deviates from the preregistered test; the deviation is noted here.

Six tests per condition: each of the three LLM dimensions (autonomy,
competence, relatedness) is compared against both non-matching
questionnaire dimensions. Across 9 conditions, this yields 54 tests
in total. All tests are two-tailed. A 95% CI on the r_within − r_cross
difference is computed via Fisher z-transformation.

```{r}
#| label: setup-h1c-helper
#| include: false

run_steiger_cond <- function(data, llm_var, q_target, q_other,
                              label_target, label_other, cond_label) {
  x  <- data[[llm_var]]
  y1 <- data[[q_target]]
  y2 <- data[[q_other]]

  complete <- complete.cases(x, y1, y2)
  x  <- x[complete]; y1 <- y1[complete]; y2 <- y2[complete]
  n  <- length(x)

  r_within <- cor(x, y1)
  r_cross  <- cor(x, y2)
  r_y1y2   <- cor(y1, y2)

  result <- cocor::cocor.dep.groups.overlap(
    r.jk = r_within, r.jh = r_cross, r.kh = r_y1y2,
    n = n, alternative = "two.sided", test = "steiger1980"
  )

  diff <- r_within - r_cross
  z1   <- 0.5 * log((1 + r_within) / (1 - r_within))
  z2   <- 0.5 * log((1 + r_cross)  / (1 - r_cross))
  se   <- sqrt(2 / (n - 3))
  crit <- qnorm(0.975)

  data.frame(
    condition  = cond_label,
    llm_dim    = llm_var,
    comparison = paste(label_target, "vs", label_other),
    r_within   = round(r_within, 2),
    r_cross    = round(r_cross,  3),
    diff       = round(diff,     3),
    ci_lo      = round(tanh((z1 - z2) - crit * se), 2),
    ci_hi      = round(tanh((z1 - z2) + crit * se), 2),
    steiger_z  = round(result@steiger1980$statistic, 2),
    p_value    = round(result@steiger1980$p.value,   3),
    n          = n
  )
}
```

```{r}
#| label: result-h1c
#| echo: false

tbl_h1c <- purrr::map_dfr(seq_len(nrow(conditions)), function(i) {
  lbl <- paste0(conditions$text_count[i], " texts × ",
                conditions$target_words[i], "w")
  sub <- filter(d_h1,
                target_words == conditions$target_words[i],
                text_count   == conditions$text_count[i])
  bind_rows(
    run_steiger_cond(sub, "llm_aut_mean", "q_aut_mean", "q_com_mean", "Q_Aut", "Q_Com", lbl),
    run_steiger_cond(sub, "llm_aut_mean", "q_aut_mean", "q_rel_mean", "Q_Aut", "Q_Rel", lbl),
    run_steiger_cond(sub, "llm_com_mean", "q_com_mean", "q_aut_mean", "Q_Com", "Q_Aut", lbl),
    run_steiger_cond(sub, "llm_com_mean", "q_com_mean", "q_rel_mean", "Q_Com", "Q_Rel", lbl),
    run_steiger_cond(sub, "llm_rel_mean", "q_rel_mean", "q_aut_mean", "Q_Rel", "Q_Aut", lbl),
    run_steiger_cond(sub, "llm_rel_mean", "q_rel_mean", "q_com_mean", "Q_Rel", "Q_Com", lbl)
  )
})

tbl_h1c |>
  flextable::flextable() |>
  style_apa7() |>
  flextable::set_caption("H1c: Discriminant Validity — Steiger's Test per Condition (N=54 tests)")
```

Two complementary figures follow. The **heatmap** (fig-h1c-heatmap) shows the full
3 × 3 LLM × Q correlation matrix for each condition; the diagonal (bold border)
= convergent validity, off-diagonal = discriminant. The **MTMM dot plot**
(fig-h1c-mtmm-dots) follows Campbell-Fiske practice: for each LLM dimension,
the filled circle (convergent r) should lie to the right of all open circles
(discriminant r) in every row; when it does, the criterion is met for that
condition.

```{r}
#| label: fig-h1c-data-prep
#| include: false

# Reconstruct 3×3 LLM × Q matrix from the 6-row-per-condition tbl_h1c
h1c_mat <- dplyr::bind_rows(
  # Diagonal: r_within. Deduplicate — same value appears in both rows per llm_dim.
  tbl_h1c |>
    dplyr::group_by(condition, llm_dim) |>
    dplyr::slice(1) |>
    dplyr::ungroup() |>
    dplyr::transmute(
      condition, llm_dim,
      q_dim   = sub("^(Q_\\w+) vs.*", "\\1", comparison),
      r       = r_within,
      is_diag = TRUE
    ),
  # Off-diagonal: r_cross
  tbl_h1c |>
    dplyr::transmute(
      condition, llm_dim,
      q_dim   = sub(".*vs (Q_\\w+)$", "\\1", comparison),
      r       = r_cross,
      is_diag = FALSE
    )
) |>
  dplyr::mutate(
    text_count   = as.integer(sub("^(\\d+) texts.*", "\\1", condition)),
    target_words = as.integer(sub(".*(\\d+)w$",      "\\1", condition)),
    llm_f  = factor(llm_dim,
                    levels = c("llm_rel_mean", "llm_com_mean", "llm_aut_mean"),
                    labels = c("LLM Rel",      "LLM Com",      "LLM Aut")),
    q_f    = factor(q_dim,
                    levels = c("Q_Aut", "Q_Com", "Q_Rel"),
                    labels = c("Q Aut", "Q Com", "Q Rel")),
    target_f = factor(target_words, levels = c(50L, 150L, 300L),
                      labels = c("50w", "150w", "300w")),
    text_f   = factor(text_count, levels = c(1L, 3L, 5L),
                      labels = c("1 text", "3 texts", "5 texts")),
    r_label  = sprintf("%.2f", r),
    txt_col  = if_else(r > 0.30, "white", "black")
  )

# Long format for MTMM dot plot
h1c_long <- dplyr::bind_rows(
  tbl_h1c |>
    dplyr::group_by(condition, llm_dim) |>
    dplyr::slice(1) |>
    dplyr::ungroup() |>
    dplyr::transmute(
      condition, llm_dim,
      q_dim  = sub("^(Q_\\w+) vs.*", "\\1", comparison),
      r_type = "Convergent",
      r      = r_within
    ),
  tbl_h1c |>
    dplyr::transmute(
      condition, llm_dim,
      q_dim  = sub(".*vs (Q_\\w+)$", "\\1", comparison),
      r_type = "Discriminant",
      r      = r_cross
    )
) |>
  dplyr::mutate(
    text_count   = as.integer(sub("^(\\d+) texts.*", "\\1", condition)),
    target_words = as.integer(sub(".*(\\d+)w$",      "\\1", condition)),
    llm_f  = factor(llm_dim,
                    levels = c("llm_aut_mean", "llm_com_mean", "llm_rel_mean"),
                    labels = c("LLM Autonomy", "LLM Competence", "LLM Relatedness")),
    q_f    = factor(q_dim,
                    levels = c("Q_Aut", "Q_Com", "Q_Rel"),
                    labels = c("Q Aut", "Q Com", "Q Rel")),
    cond_f = factor(
      paste0(text_count, "t × ", target_words, "w"),
      levels = paste0(rep(c(1L, 3L, 5L), each = 3L), "t × ",
                      rep(c(50L, 150L, 300L), 3L), "w")
    )
  )
```

```{r}
#| label: fig-h1c-heatmap
#| echo: false
#| fig-cap: |
#|   _H1c: LLM × Questionnaire Correlation Matrix per Condition (Pearson r)._
#|   _Diagonal (bold border) = convergent validity (same need, different method);_
#|   _off-diagonal = discriminant validity. Colour intensity = r magnitude._
#|   _Rows = number of texts; columns = target word length._
#| fig-width: 7
#| fig-height: 7
#| fig-dpi: 300

ggplot(h1c_mat, aes(x = q_f, y = llm_f, fill = r)) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_tile(data = dplyr::filter(h1c_mat, is_diag),
            color = "black", linewidth = 1.1, fill = NA) +
  geom_text(aes(label = r_label, color = txt_col),
            size = 2.2, family = "Calibri") +
  scale_color_identity(guide = "none") +
  scale_fill_gradient(low = "#EAF4FB", high = "#1F5A71",
                      name = "r", limits = c(0, 1)) +
  facet_grid(text_f ~ target_f) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 9) +
  theme(
    text             = element_text(family = "Calibri"),
    strip.text       = element_text(face = "plain", colour = "black", size = 9),
    panel.grid       = element_blank(),
    axis.text.x      = element_text(size = 8, angle = 30, hjust = 1),
    axis.text.y      = element_text(size = 8),
    legend.position  = "right",
    panel.spacing    = unit(0.6, "lines")
  )
```

```{r}
#| label: fig-h1c-mtmm-dots
#| echo: false
#| fig-cap: |
#|   _H1c: MTMM Validity Coefficient Comparison (Campbell-Fiske criterion)._
#|   _Each panel = one LLM dimension; y-axis = condition; x-axis = Pearson r._
#|   _Filled circle = convergent r (same need); open circles = discriminant r_
#|   _(non-matching Q dimensions, coloured by Q dim). Discriminant validity holds_
#|   _when the filled circle lies to the right of all open circles in a row._
#| fig-width: 10
#| fig-height: 6
#| fig-dpi: 500

ggplot(h1c_long,
       aes(x = r, y = cond_f,
           group = interaction(r_type, q_f),
           shape = r_type, color = q_f, size = r_type)) +
  geom_vline(xintercept = 0, color = "grey70", linewidth = 0.3) +
  geom_point(position = position_dodge(width = 0.6), alpha = 0.9) +
  scale_shape_manual(values = c("Convergent" = 19L, "Discriminant" = 1L),
                     name = NULL) +
  scale_size_manual(values  = c("Convergent" = 3.5, "Discriminant" = 2.5),
                    guide   = "none") +
  scale_color_manual(
    values = c("Q Aut" = "#1F5A71", "Q Com" = "#2085AB", "Q Rel" = "#6DBCE0"),
    name   = "Q dimension"
  ) +
  facet_wrap(~llm_f, ncol = 3L) +
  labs(x = "Correlation (r)", y = NULL) +
  theme_minimal(base_size = 11) +
  theme(
    text               = element_text(family = "Calibri"),
    strip.text         = element_text(face = "plain", colour = "black", size = 11),
    panel.grid.minor   = element_blank(),
    panel.grid.major.y = element_line(linewidth = 0.2, color = "grey88"),
    legend.position    = "top",
    axis.title.x       = element_text(margin = margin(t = 10))
  )
```

# 4 H1d: Structural Consistency

Structural consistency asks whether the inter-need correlation pattern
in LLM scores mirrors the pattern in questionnaire scores. Three
inter-need correlations are computed for each source (Q and LLM): Aut–Com,
Aut–Rel, Com–Rel. The Pearson correlation between these two 3-element
vectors is the meta-correlation (meta-r).

Because N = 3 data points, no inferential test is applied. Results
are descriptive only. Note that questionnaire inter-need correlations
are constant across conditions (Q scores do not vary by condition);
any condition-to-condition variation in meta-r reflects shifts in the
LLM's inter-need structure.

```{r}
#| label: result-h1d
#| echo: false

tbl_h1d <- purrr::map_dfr(seq_len(nrow(conditions)), function(i) {
  lbl <- paste0(conditions$text_count[i], " texts × ",
                conditions$target_words[i], "w")
  sub <- filter(d_h1,
                target_words == conditions$target_words[i],
                text_count   == conditions$text_count[i])

  q_vec <- c(
    cor(sub$q_aut_mean, sub$q_com_mean, use = "complete.obs"),
    cor(sub$q_aut_mean, sub$q_rel_mean, use = "complete.obs"),
    cor(sub$q_com_mean, sub$q_rel_mean, use = "complete.obs")
  )
  llm_vec <- c(
    cor(sub$llm_aut_mean, sub$llm_com_mean, use = "complete.obs"),
    cor(sub$llm_aut_mean, sub$llm_rel_mean, use = "complete.obs"),
    cor(sub$llm_com_mean, sub$llm_rel_mean, use = "complete.obs")
  )

  data.frame(
    condition = lbl,
    pair      = c("Aut-Com", "Aut-Rel", "Com-Rel"),
    r_Q       = round(q_vec,   3),
    r_LLM     = round(llm_vec, 2),
    meta_r    = round(cor(q_vec, llm_vec), 2)
  )
})

tbl_h1d |>
  flextable::flextable() |>
  style_apa7() |>
  flextable::set_caption("H1d: Structural Consistency — Inter-need correlations and meta-r per condition (N=3 pairs, descriptive only)")
```

```{r}
#| label: fig-h1d-structural-consistency
#| echo: false
#| fig-cap: |
#|   _H1d: LLM Inter-Need Correlations vs. Questionnaire Reference (Pearson r)._
#|   _Each panel shows one need pair. Lines and points = LLM r per condition,_
#|   _colored by number of texts (light to dark = 1, 3, 5). Dashed line = constant_
#|   _questionnaire r. Convergence toward the dashed line indicates structural consistency._
#| fig-width: 9
#| fig-height: 4
#| fig-dpi: 300

COL_TEXTS_H1 <- c("1" = "#6DBCE0", "3" = "#2085AB", "5" = "#1F5A71")

fig_h1d_data <- tbl_h1d |>
  mutate(
    text_count   = as.integer(sub("^(\\d+) texts.*", "\\1", condition)),
    target_words = as.integer(sub(".*(\\d+)w$",      "\\1", condition)),
    pair_f = factor(pair,
                    levels = c("Aut-Com", "Aut-Rel", "Com-Rel"),
                    labels = c("Autonomy–Competence",
                               "Autonomy–Relatedness",
                               "Competence–Relatedness"))
  )

q_refs <- fig_h1d_data |>
  distinct(pair_f, r_Q)

ggplot(fig_h1d_data,
       aes(x = target_words, y = r_LLM,
           color = factor(text_count), group = factor(text_count))) +
  geom_hline(data = q_refs, aes(yintercept = r_Q),
             linetype = "dashed", color = "grey40", linewidth = 0.5,
             inherit.aes = FALSE) +
  geom_text(data = q_refs, aes(x = 50, y = r_Q, label = sprintf("Q = %.2f", r_Q)),
            hjust = 0, vjust = -0.4, size = 2.8, color = "grey40",
            family = "Calibri", inherit.aes = FALSE) +
  geom_line(linewidth = 0.55,
            position = position_dodge(width = 15)) +
  geom_point(size = 2.5,
             position = position_dodge(width = 15)) +
  facet_wrap(~pair_f, ncol = 3L) +
  scale_x_continuous(breaks = c(50L, 150L, 300L),
                     labels = c("50w", "150w", "300w")) +
  scale_color_manual(values = COL_TEXTS_H1, name = "Number of texts") +
  labs(x = "Target word length", y = "Inter-need correlation (r)") +
  theme_minimal(base_size = 11) +
  theme(
    text               = element_text(family = "Calibri"),
    strip.text         = element_text(face = "plain", colour = "black", size = 11),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    legend.position    = "top",
    axis.title.x       = element_text(margin = margin(t = 10)),
    axis.title.y       = element_text(margin = margin(r = 10))
  )
```

# 5 Spearman Robustness Checks

## 5.1 H1c (Spearman): Discriminant Validity

Same analysis as Section 3 with Spearman ρ substituted for Pearson r in
all three correlation inputs (`r_within`, `r_cross`, `r_y1y2`). Steiger's
(1980) test is designed for Pearson r; applying it to Spearman ρ is an
approximation and results should be interpreted accordingly.

```{r}
#| label: setup-h1c-spearman-helper
#| include: false

run_steiger_cond_spearman <- function(data, llm_var, q_target, q_other,
                                       label_target, label_other, cond_label) {
  x  <- data[[llm_var]]
  y1 <- data[[q_target]]
  y2 <- data[[q_other]]

  complete <- complete.cases(x, y1, y2)
  x  <- x[complete]; y1 <- y1[complete]; y2 <- y2[complete]
  n  <- length(x)

  r_within <- cor(x, y1, method = "spearman")
  r_cross  <- cor(x, y2, method = "spearman")
  r_y1y2   <- cor(y1, y2, method = "spearman")

  result <- cocor::cocor.dep.groups.overlap(
    r.jk = r_within, r.jh = r_cross, r.kh = r_y1y2,
    n = n, alternative = "two.sided", test = "steiger1980"
  )

  diff <- r_within - r_cross
  z1   <- 0.5 * log((1 + r_within) / (1 - r_within))
  z2   <- 0.5 * log((1 + r_cross)  / (1 - r_cross))
  se   <- sqrt(2 / (n - 3))
  crit <- qnorm(0.975)

  data.frame(
    condition  = cond_label,
    llm_dim    = llm_var,
    comparison = paste(label_target, "vs", label_other),
    rho_within = round(r_within, 2),
    rho_cross  = round(r_cross,  3),
    diff       = round(diff,     3),
    ci_lo      = round(tanh((z1 - z2) - crit * se), 2),
    ci_hi      = round(tanh((z1 - z2) + crit * se), 2),
    steiger_z  = round(result@steiger1980$statistic, 2),
    p_value    = round(result@steiger1980$p.value,   3),
    n          = n
  )
}
```

```{r}
#| label: result-h1c-spearman
#| echo: false

tbl_h1c_spearman <- purrr::map_dfr(seq_len(nrow(conditions)), function(i) {
  lbl <- paste0(conditions$text_count[i], " texts × ",
                conditions$target_words[i], "w")
  sub <- filter(d_h1,
                target_words == conditions$target_words[i],
                text_count   == conditions$text_count[i])
  bind_rows(
    run_steiger_cond_spearman(sub, "llm_aut_mean", "q_aut_mean", "q_com_mean", "Q_Aut", "Q_Com", lbl),
    run_steiger_cond_spearman(sub, "llm_aut_mean", "q_aut_mean", "q_rel_mean", "Q_Aut", "Q_Rel", lbl),
    run_steiger_cond_spearman(sub, "llm_com_mean", "q_com_mean", "q_aut_mean", "Q_Com", "Q_Aut", lbl),
    run_steiger_cond_spearman(sub, "llm_com_mean", "q_com_mean", "q_rel_mean", "Q_Com", "Q_Rel", lbl),
    run_steiger_cond_spearman(sub, "llm_rel_mean", "q_rel_mean", "q_aut_mean", "Q_Rel", "Q_Aut", lbl),
    run_steiger_cond_spearman(sub, "llm_rel_mean", "q_rel_mean", "q_com_mean", "Q_Rel", "Q_Com", lbl)
  )
})

tbl_h1c_spearman |>
  flextable::flextable() |>
  style_apa7() |>
  flextable::set_caption("H1c (Spearman): Discriminant Validity — Steiger's Test per Condition (N=54 tests; ρ as approximation)")
```

## 5.2 H1d (Spearman): Structural Consistency

Same analysis as Section 4 with Spearman ρ substituted for Pearson r for
all six inter-need correlations. The meta-r between the two 3-element
vectors remains Pearson (rank-ordering 3 values is not meaningful).

```{r}
#| label: result-h1d-spearman
#| echo: false

tbl_h1d_spearman <- purrr::map_dfr(seq_len(nrow(conditions)), function(i) {
  lbl <- paste0(conditions$text_count[i], " texts × ",
                conditions$target_words[i], "w")
  sub <- filter(d_h1,
                target_words == conditions$target_words[i],
                text_count   == conditions$text_count[i])

  q_vec <- c(
    cor(sub$q_aut_mean, sub$q_com_mean, use = "complete.obs", method = "spearman"),
    cor(sub$q_aut_mean, sub$q_rel_mean, use = "complete.obs", method = "spearman"),
    cor(sub$q_com_mean, sub$q_rel_mean, use = "complete.obs", method = "spearman")
  )
  llm_vec <- c(
    cor(sub$llm_aut_mean, sub$llm_com_mean, use = "complete.obs", method = "spearman"),
    cor(sub$llm_aut_mean, sub$llm_rel_mean, use = "complete.obs", method = "spearman"),
    cor(sub$llm_com_mean, sub$llm_rel_mean, use = "complete.obs", method = "spearman")
  )

  data.frame(
    condition = lbl,
    pair      = c("Aut-Com", "Aut-Rel", "Com-Rel"),
    rho_Q     = round(q_vec,   3),
    rho_LLM   = round(llm_vec, 2),
    meta_r    = round(cor(q_vec, llm_vec), 2)
  )
})

tbl_h1d_spearman |>
  flextable::flextable() |>
  style_apa7() |>
  flextable::set_caption("H1d (Spearman): Structural Consistency — Inter-need ρ and meta-r per condition (N=3 pairs, descriptive only)")
```

# 6 Session Info

```{r}
#| label: session-info
#| include: false

sessionInfo()
```
