---
title: "RQ3 Data — Human Import, Merge, and Descriptives"
subtitle: "Human participant data · AI–human merge · Matched condition"
author: "Caroline Müller"
date: "2026-05-17"
format:
  html:
    code-fold: true
    code-summary: "Show code"
    code-tools: true
    toc: true
    self-contained: true
    fig-cap-location: top
editor:
  markdown:
    wrap: 72
knitr:
  opts_chunk:
    dpi: 300
---

# 1 Overview

**Prerequisite:** Run `10_RQ2_data.qmd` first to generate
`d_rq3_ai.rds`, `tbl_icc_raw.rds`, and
`tbl_word_count_by_condition.rds` in `10_data/80_analyses_exports/`.

This file covers:

- Import and quality-check of human participant CSV
- Questionnaire composites (autonomy: 4 items; competence: 4 items;
  relatedness: **5 items** — differs from AI surrogates)
- Merge with matched AI surrogate condition (short-3: target_words = 50,
  text_count = 3)
- Descriptive comparison (demographics, Q scores, LLM scores)
- Export of `d_rq3_merged.rds` and `d_human_clean.rds`

**Note on relatedness items:** human participants completed 5 relatedness
items (rel1–rel5); AI surrogate composites use 4 items. The composite
means (`q_rel_mean`) are therefore not strictly comparable at the item
level, but are on the same 1–5 scale. This is acknowledged in all
comparisons.

---

# 2 Setup

## 2.1 Environment

```{r}
#| label: setup-environment
#| include: false
#| warning: false

graphics.off()

suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(tidyr)
  library(flextable)
  library(scales)
  library(psych)
  library(officer)
  library(janitor)
  library(systemfonts)
  library(ragg)
  library(stringr)
})

knitr::opts_chunk$set(dev = "ragg_png")
if (.Platform$OS.type == "windows") windowsFonts(Calibri = windowsFont("Calibri"))

message("Libraries loaded.")
```

## 2.2 Shared Constants

```{r}
#| label: setup-common
#| include: false

source("00_common.R")
```

## 2.3 Load Upstream Exports

```{r}
#| label: setup-imports
#| include: false

EXPORT_DIR   <- "../10_data/80_analyses_exports"
RAW_DATA_DIR <- "../10_data/10_raw"

rds_paths <- list(
  ai_rq3 = file.path(EXPORT_DIR, "d_rq3_ai.rds")
)

for (nm in names(rds_paths)) {
  if (!file.exists(rds_paths[[nm]])) {
    stop("Missing upstream export — run 10_RQ2_data.qmd first: ", rds_paths[[nm]])
  }
}

d_rq3_ai <- readRDS(rds_paths$ai_rq3)

message("AI data loaded: ", nrow(d_rq3_ai), " rows, ",
        n_distinct(d_rq3_ai$id), " AI surrogates.")
```

---

# 3 Human Data Import

## 3.1 Import Raw CSV

```{r}
#| label: import-human-raw
#| include: false

HUMAN_CSV <- file.path(RAW_DATA_DIR, "humans_2026-04-13.csv")

if (!file.exists(HUMAN_CSV))
  stop("Human data file not found: ", HUMAN_CSV)

raw_human <- read.csv(HUMAN_CSV, sep = "|", na.strings = "NA",
                      stringsAsFactors = FALSE)

message("Human CSV loaded: ", basename(HUMAN_CSV),
        " | Modified: ", format(file.mtime(HUMAN_CSV), "%Y-%m-%d %H:%M"),
        " | Rows: ", nrow(raw_human),
        " | Columns: ", ncol(raw_human))
```

## 3.2 Import Audit

```{r}
#| label: tbl-import-audit-human
#| echo: false
#| tbl-cap: "Human Data Import Audit"

tbl_import_audit_human <- tibble::tribble(
  ~Metric,                ~Value,
  "Source file",          basename(HUMAN_CSV),
  "Total rows",           as.character(nrow(raw_human)),
  "Expected columns",     "participant_id, gender, birthyear, education, aut1-4, comp1-4, rel1-5, PVQ_M/K_1-21, llm_*_rep1-3, wc_*",
  "Actual columns",       as.character(ncol(raw_human)),
  "Q aut items",          if (all(paste0("aut", 1:4) %in% names(raw_human))) "aut1-4 present" else "MISSING",
  "Q com items",          if (all(paste0("comp", 1:4) %in% names(raw_human))) "comp1-4 present" else "MISSING",
  "Q rel items",          if (all(paste0("rel", 1:5) %in% names(raw_human))) "rel1-5 present (5 items)" else "MISSING",
  "LLM aut columns",      if (all(paste0("llm_aut_rep", 1:3) %in% names(raw_human))) "llm_aut_rep1-3 present" else "MISSING",
  "LLM com columns",      if (all(paste0("llm_com_rep", 1:3) %in% names(raw_human))) "llm_com_rep1-3 present" else "MISSING",
  "LLM rel columns",      if (all(paste0("llm_rel_rep", 1:3) %in% names(raw_human))) "llm_rel_rep1-3 present" else "MISSING",
  "Word count col",       if ("wc_total" %in% names(raw_human)) "wc_total present" else "MISSING",
  "Duplicate IDs",        as.character(sum(duplicated(raw_human$participant_id)))
)

tbl_import_audit_human |>
  flextable() |>
  style_apa7() |>
  add_apa_note("Column presence gates — any MISSING entry would indicate a pipeline change in the source file.")
```

## 3.3 Data Quality Checks

```{r}
#| label: check-human-quality
#| include: false

# ── Range checks ─────────────────────────────────────────────────────────────

q_items_raw  <- raw_human |> dplyr::select(aut1:aut4, comp1:comp4, rel1:rel5)
llm_items_raw <- raw_human |> dplyr::select(dplyr::starts_with("llm_"))

check_human_q_range <- q_items_raw |>
  tidyr::pivot_longer(everything(), names_to = "item", values_to = "val") |>
  dplyr::filter(!is.na(val)) |>
  dplyr::summarise(
    min_val = min(val), max_val = max(val),
    n_below_1 = sum(val < 1), n_above_5 = sum(val > 5)
  )

check_human_llm_range <- llm_items_raw |>
  tidyr::pivot_longer(everything(), names_to = "item", values_to = "val") |>
  dplyr::filter(!is.na(val)) |>
  dplyr::summarise(
    min_val = min(val), max_val = max(val),
    n_below_1 = sum(val < 1), n_above_5 = sum(val > 5)
  )

if (check_human_q_range$n_below_1 > 0 | check_human_q_range$n_above_5 > 0)
  warning("Q items out of [1, 5] range: below=", check_human_q_range$n_below_1,
          " above=", check_human_q_range$n_above_5)

if (check_human_llm_range$n_below_1 > 0 | check_human_llm_range$n_above_5 > 0)
  warning("LLM items out of [1, 5] range: below=", check_human_llm_range$n_below_1,
          " above=", check_human_llm_range$n_above_5)

cat("Q item range: [", check_human_q_range$min_val, ",",
    check_human_q_range$max_val, "] —",
    ifelse(check_human_q_range$n_below_1 + check_human_q_range$n_above_5 == 0,
           "all in [1, 5].", "OUT-OF-RANGE VALUES DETECTED."), "\n")

cat("LLM item range: [", check_human_llm_range$min_val, ",",
    check_human_llm_range$max_val, "] —",
    ifelse(check_human_llm_range$n_below_1 + check_human_llm_range$n_above_5 == 0,
           "all in [1, 5].", "OUT-OF-RANGE VALUES DETECTED."), "\n")

# ── Birthyear plausibility ────────────────────────────────────────────────────

current_year <- 2026L
check_human_birthyear <- raw_human |>
  dplyr::mutate(
    age_computed = current_year - as.integer(birthyear),
    implausible  = age_computed < 18 | age_computed > 80
  ) |>
  dplyr::summarise(
    n_total      = dplyr::n(),
    n_implausible = sum(implausible, na.rm = TRUE),
    min_age      = min(age_computed, na.rm = TRUE),
    max_age      = max(age_computed, na.rm = TRUE),
    mean_age     = round(mean(age_computed, na.rm = TRUE), 1)
  )

cat("Birthyear → age range: [", check_human_birthyear$min_age, ",",
    check_human_birthyear$max_age, "], M =", check_human_birthyear$mean_age, "\n")

if (check_human_birthyear$n_implausible > 0)
  warning("Implausible ages (< 18 or > 80): n = ",
          check_human_birthyear$n_implausible)

# ── Missing data ─────────────────────────────────────────────────────────────

check_human_missing <- colSums(is.na(raw_human)) |>
  tibble::enframe(name = "column", value = "n_missing") |>
  dplyr::filter(n_missing > 0) |>
  dplyr::arrange(dplyr::desc(n_missing))

if (nrow(check_human_missing) > 0) {
  cat("Columns with missing data:\n")
  print(check_human_missing)
} else {
  cat("Missing data check OK: no NAs in any column.\n")
}

# ── Duplicate IDs ─────────────────────────────────────────────────────────────

check_human_dup_ids <- sum(duplicated(raw_human$participant_id))
if (check_human_dup_ids > 0) {
  stop("CRITICAL: ", check_human_dup_ids, " duplicate participant_id(s) found.")
} else {
  cat("Duplicate ID check OK: all participant_ids unique.\n")
}
```

## 3.4 Type Conversion, Rename, Drop PVQ

```{r}
#| label: clean-human
#| include: false

# ── All columns named explicitly — upstream rename breaks loudly here ─────────

d_human_precomp <- raw_human |>
  dplyr::rename(
    # Demographics
    id        = participant_id,
    gender    = gender,
    birthyear = birthyear,
    edu_level = education,
    # Questionnaire items (aut/com/rel → q_aut/q_com/q_rel naming; comp→com)
    q_aut_1 = aut1,  q_aut_2 = aut2,  q_aut_3 = aut3,  q_aut_4 = aut4,
    q_com_1 = comp1, q_com_2 = comp2, q_com_3 = comp3, q_com_4 = comp4,
    q_rel_1 = rel1,  q_rel_2 = rel2,  q_rel_3 = rel3,  q_rel_4 = rel4,
    q_rel_5 = rel5,
    # LLM repeats
    llm_aut_1 = llm_aut_rep1, llm_aut_2 = llm_aut_rep2, llm_aut_3 = llm_aut_rep3,
    llm_com_1 = llm_com_rep1, llm_com_2 = llm_com_rep2, llm_com_3 = llm_com_rep3,
    llm_rel_1 = llm_rel_rep1, llm_rel_2 = llm_rel_rep2, llm_rel_3 = llm_rel_rep3,
    # Word counts
    actual_words = wc_total,
    wc_text_1    = wc_text_1,
    wc_text_2    = wc_text_2,
    wc_text_3    = wc_text_3
  ) |>
  dplyr::select(-dplyr::starts_with("PVQ")) |>
  dplyr::mutate(
    id        = as.character(id),
    gender    = as.integer(gender),
    edu_level = as.integer(edu_level),
    birthyear = as.integer(birthyear),
    dplyr::across(dplyr::starts_with("q_"),   as.numeric),
    dplyr::across(dplyr::starts_with("llm_"), as.numeric),
    actual_words = as.integer(actual_words),
    wc_text_1    = as.integer(wc_text_1),
    wc_text_2    = as.integer(wc_text_2),
    wc_text_3    = as.integer(wc_text_3)
  )

message("Human data cleaned: ", nrow(d_human_precomp), " rows, ",
        ncol(d_human_precomp), " columns.")
```

## 3.5 Age Computation

```{r}
#| label: compute-age
#| include: false

STUDY_YEAR <- 2026L

d_human_precomp <- d_human_precomp |>
  dplyr::mutate(age_year = STUDY_YEAR - birthyear)

stopifnot(
  "age_year: all in plausible range 18–80" =
    all(d_human_precomp$age_year >= 18 & d_human_precomp$age_year <= 80, na.rm = TRUE)
)

message("Age computed: M = ", round(mean(d_human_precomp$age_year), 1),
        ", range [", min(d_human_precomp$age_year), ", ",
        max(d_human_precomp$age_year), "]")
```

## 3.6 Questionnaire Composites

**Item counts:** autonomy = 4 items, competence = 4 items, relatedness
= **5 items** (rel1–rel5). The relatedness composite for human
participants therefore averages 5 items, compared to 4 for AI
surrogates. Both are on the same 1–5 scale.

```{r}
#| label: compute-q-composites-human
#| include: false

d_human_precomp <- d_human_precomp |>
  dplyr::mutate(
    # 4 items each (same as AI surrogates)
    q_aut_mean = rowMeans(across(c(q_aut_1, q_aut_2, q_aut_3, q_aut_4)),
                          na.rm = TRUE),
    q_com_mean = rowMeans(across(c(q_com_1, q_com_2, q_com_3, q_com_4)),
                          na.rm = TRUE),
    # 5 items for humans (rel_5 is the additional item vs AI's 4)
    q_rel_mean = rowMeans(across(c(q_rel_1, q_rel_2, q_rel_3, q_rel_4, q_rel_5)),
                          na.rm = TRUE)
  )

stopifnot(
  "q_aut_mean in [1, 5]" = all(dplyr::between(d_human_precomp$q_aut_mean, 1, 5), na.rm = TRUE),
  "q_com_mean in [1, 5]" = all(dplyr::between(d_human_precomp$q_com_mean, 1, 5), na.rm = TRUE),
  "q_rel_mean in [1, 5]" = all(dplyr::between(d_human_precomp$q_rel_mean, 1, 5), na.rm = TRUE)
)

message("Q composites computed (4 aut, 4 com, 5 rel items).")
```

## 3.7 LLM Composite Means

```{r}
#| label: compute-llm-composites-human
#| include: false

d_human_precomp <- d_human_precomp |>
  dplyr::mutate(
    llm_aut_mean = rowMeans(across(c(llm_aut_1, llm_aut_2, llm_aut_3)), na.rm = TRUE),
    llm_com_mean = rowMeans(across(c(llm_com_1, llm_com_2, llm_com_3)), na.rm = TRUE),
    llm_rel_mean = rowMeans(across(c(llm_rel_1, llm_rel_2, llm_rel_3)), na.rm = TRUE)
  )

stopifnot(
  "llm_aut_mean in [1, 5]" = all(dplyr::between(d_human_precomp$llm_aut_mean, 1, 5), na.rm = TRUE),
  "llm_com_mean in [1, 5]" = all(dplyr::between(d_human_precomp$llm_com_mean, 1, 5), na.rm = TRUE),
  "llm_rel_mean in [1, 5]" = all(dplyr::between(d_human_precomp$llm_rel_mean, 1, 5), na.rm = TRUE)
)

# Final clean human dataset (all item-level columns preserved for d_human_clean)
d_human_clean <- d_human_precomp |>
  dplyr::mutate(id_type = "HU",
                condition    = "short-3",
                text_count   = 3L,
                target_words = 50L)

message("LLM composites computed. d_human_clean: ", nrow(d_human_clean),
        " human participants.")
```

---

# 4 Merge AI + Human

## 4.1 Matched AI Condition

`d_rq3_ai.rds` was filtered to short-3 (target_words = 50, text_count
= 3) in `10_RQ2_data.qmd`. Verify the condition before merging.

```{r}
#| label: verify-ai-condition
#| include: false

stopifnot(
  "d_rq3_ai: only short-3 condition" =
    all(d_rq3_ai$condition == "short-3"),
  "d_rq3_ai: target_words = 50" =
    all(d_rq3_ai$target_words == 50L),
  "d_rq3_ai: text_count = 3" =
    all(d_rq3_ai$text_count == 3L),
  "d_rq3_ai: id_type = AI" =
    all(d_rq3_ai$id_type == "AI")
)

message("AI condition verified: short-3 (target_words=50, text_count=3), N = ",
        n_distinct(d_rq3_ai$id), " surrogates.")
```

## 4.2 Merge

```{r}
#| label: merge-ai-human
#| include: false

# Select only the columns needed for RQ3 analyses (+ condition metadata)
# text_count is factor in d_rq3_ai (from 10_RQ2_data pipeline) but integer in
# d_human_clean; coerce via character to preserve the numeric value, not the level index
d_ai_for_merge <- d_rq3_ai |>
  dplyr::mutate(text_count = as.integer(as.character(text_count))) |>
  dplyr::select(id, id_type, llm_aut_mean, llm_com_mean, llm_rel_mean,
                q_aut_mean, q_com_mean, q_rel_mean, actual_words,
                condition, text_count, target_words)

d_hu_for_merge <- d_human_clean |>
  dplyr::select(id, id_type, llm_aut_mean, llm_com_mean, llm_rel_mean,
                q_aut_mean, q_com_mean, q_rel_mean, actual_words,
                condition, text_count, target_words)

d_rq3_merged <- dplyr::bind_rows(d_ai_for_merge, d_hu_for_merge)

stopifnot(
  "merged: rows = AI + human" =
    nrow(d_rq3_merged) == nrow(d_ai_for_merge) + nrow(d_hu_for_merge),
  "merged: exactly two id_type groups" =
    n_distinct(d_rq3_merged$id_type) == 2L,
  "merged: AI group present" =
    "AI" %in% d_rq3_merged$id_type,
  "merged: HU group present" =
    "HU" %in% d_rq3_merged$id_type,
  "merged: no missing LLM means" =
    !anyNA(d_rq3_merged[, c("llm_aut_mean", "llm_com_mean", "llm_rel_mean")]),
  "merged: no missing Q means" =
    !anyNA(d_rq3_merged[, c("q_aut_mean", "q_com_mean", "q_rel_mean")])
)

N_AI <- sum(d_rq3_merged$id_type == "AI")
N_HU <- sum(d_rq3_merged$id_type == "HU")

message("Merge complete: ", N_AI, " AI + ", N_HU, " human = ",
        nrow(d_rq3_merged), " total rows.")
```

## 4.3 Word Count Match Verification

```{r}
#| label: tbl-wordcount-match
#| echo: false
#| tbl-cap: "Word Count Comparison: AI Surrogates (Short-3) vs Human Participants"

d_rq3_merged |>
  dplyr::group_by(id_type) |>
  dplyr::summarise(
    N   = dplyr::n(),
    M   = round(mean(actual_words, na.rm = TRUE), 1),
    SD  = round(sd(actual_words,   na.rm = TRUE), 1),
    Mdn = round(median(actual_words, na.rm = TRUE), 0),
    Min = min(actual_words, na.rm = TRUE),
    Max = max(actual_words, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::mutate(id_type = dplyr::recode(id_type, "AI" = LABEL_AI, "HU" = LABEL_HU)) |>
  dplyr::rename(!!LABEL_SAMPLE := id_type) |>
  flextable() |>
  style_apa7() |>
  add_apa_note(paste0(
    "Actual words = total word count across all 3 scored texts. ",
    "AI surrogates: short-3 condition (target = 50 words per text, 3 texts). ",
    "Human participants wrote 3 texts targeting ~50 words each."
  ))
```

---

# 5 Descriptives: Demographics (Human)

```{r}
#| label: tbl-demo-human
#| echo: false
#| tbl-cap: "Human Participant Demographics"

# Gender labels: 1 = male, 2 = female, 3 = other/non-binary (verify against survey)
gender_labels <- c("1" = "Male", "2" = "Female", "3" = "Other / non-binary")

gender_tbl <- d_human_clean |>
  dplyr::count(gender) |>
  dplyr::mutate(
    gender_label = dplyr::recode(as.character(gender),
                                 !!!gender_labels, .default = paste("Code", gender)),
    pct = round(100 * n / sum(n), 1),
    gender_str = paste0(gender_label, ": n = ", n, " (", pct, "%)")
  ) |>
  dplyr::pull(gender_str) |>
  paste(collapse = "; ")

# Education levels (1–7+ scale; labels should be verified against survey codebook)
edu_tbl <- d_human_clean |>
  dplyr::count(edu_level) |>
  dplyr::arrange(edu_level) |>
  dplyr::mutate(
    pct = round(100 * n / sum(n), 1),
    edu_str = paste0("Level ", edu_level, ": n = ", n, " (", pct, "%)")
  ) |>
  dplyr::pull(edu_str) |>
  paste(collapse = "; ")

tibble::tribble(
  ~Characteristic,            ~Value,
  "N",                        as.character(nrow(d_human_clean)),
  "Age: M (SD)",              sprintf("%.1f (%.1f)", mean(d_human_clean$age_year),
                                      sd(d_human_clean$age_year)),
  "Age: range",               sprintf("[%d, %d]", min(d_human_clean$age_year),
                                      max(d_human_clean$age_year)),
  "Gender distribution",      gender_tbl,
  "Education distribution",   edu_tbl
) |>
  flextable() |>
  style_apa7() |>
  add_apa_note(paste0(
    "Age computed as 2026 − birth year. Gender and education labels based on ",
    "survey coding scheme — verify against original questionnaire codebook. ",
    "Education scale: 1 = some high school ... 7 = doctoral degree (verify)."
  ))
```

---

# 6 Descriptives: Questionnaire Scores

## 6.1 Summary Table by Group

```{r}
#| label: tbl-q-desc-bygroup
#| echo: false
#| tbl-cap: "Questionnaire Score Descriptives by Group and Need Dimension"

d_rq3_merged |>
  tidyr::pivot_longer(
    cols = c(q_aut_mean, q_com_mean, q_rel_mean),
    names_to = "need_var", values_to = "score"
  ) |>
  dplyr::mutate(
    Need = dplyr::case_when(
      need_var == "q_aut_mean" ~ TABLE_LABEL_AUT,
      need_var == "q_com_mean" ~ TABLE_LABEL_COM,
      need_var == "q_rel_mean" ~ TABLE_LABEL_REL
    ),
    Sample = dplyr::recode(id_type, "AI" = LABEL_AI, "HU" = LABEL_HU)
  ) |>
  dplyr::group_by(Need, Sample) |>
  dplyr::summarise(
    n   = dplyr::n(),
    M   = round(mean(score, na.rm = TRUE), 2),
    SD  = round(sd(score, na.rm = TRUE),   2),
    Mdn = round(median(score, na.rm = TRUE), 2),
    Min = round(min(score, na.rm = TRUE),   2),
    Max = round(max(score, na.rm = TRUE),   2),
    Skew = round(psych::skew(score, na.rm = TRUE), 2),
    Kurt = round(psych::kurtosi(score, na.rm = TRUE), 2),
    .groups = "drop"
  ) |>
  dplyr::arrange(Need, Sample) |>
  flextable() |>
  flextable::merge_v(j = "Need") |>
  style_apa7() |>
  add_apa_note(paste0(
    "All scores on a 1–5 scale. Skew = Fisher's skewness; Kurt = excess kurtosis. ",
    "Human q_rel_mean averages 5 items; AI q_rel_mean averages 4 items."
  ))
```

## 6.2 Distribution by Group

```{r}
#| label: fig-q-dist-bygroup
#| echo: false
#| fig-cap: "Questionnaire Score Distributions by Group and Need Dimension. Violin + boxplot. AI surrogates (blue) vs human participants (red)."
#| fig-width: 9
#| fig-height: 4
#| fig-dpi: 300

d_rq3_merged |>
  tidyr::pivot_longer(
    cols = c(q_aut_mean, q_com_mean, q_rel_mean),
    names_to = "need_var", values_to = "score"
  ) |>
  dplyr::mutate(
    Need   = factor(dplyr::case_when(
      need_var == "q_aut_mean" ~ TABLE_LABEL_AUT,
      need_var == "q_com_mean" ~ TABLE_LABEL_COM,
      need_var == "q_rel_mean" ~ TABLE_LABEL_REL
    ), levels = NEEDS),
    Sample = dplyr::recode(id_type, "AI" = LABEL_AI, "HU" = LABEL_HU)
  ) |>
  ggplot(aes(x = Sample, y = score, fill = id_type)) +
  geom_violin(alpha = 0.4, trim = FALSE, linewidth = 0.4) +
  geom_boxplot(width = 0.18, outlier.size = 0.8, outlier.alpha = 0.5) +
  scale_fill_manual(values = c("AI" = COL_AI_SINGLE, "HU" = COL_HU_SINGLE),
                    guide = "none") +
  scale_y_continuous(limits = c(1, 5), breaks = 1:5) +
  facet_wrap(~Need, ncol = 3L) +
  labs(x = NULL, y = "Questionnaire score") +
  theme_thesis()
```

---

# 7 Descriptives: LLM Scores

## 7.1 Summary Table by Group

```{r}
#| label: tbl-llm-desc-bygroup
#| echo: false
#| tbl-cap: "LLM Score Descriptives by Group and Need Dimension"

d_rq3_merged |>
  tidyr::pivot_longer(
    cols = c(llm_aut_mean, llm_com_mean, llm_rel_mean),
    names_to = "need_var", values_to = "score"
  ) |>
  dplyr::mutate(
    Need = dplyr::case_when(
      need_var == "llm_aut_mean" ~ TABLE_LABEL_AUT,
      need_var == "llm_com_mean" ~ TABLE_LABEL_COM,
      need_var == "llm_rel_mean" ~ TABLE_LABEL_REL
    ),
    Sample = dplyr::recode(id_type, "AI" = LABEL_AI, "HU" = LABEL_HU)
  ) |>
  dplyr::group_by(Need, Sample) |>
  dplyr::summarise(
    n   = dplyr::n(),
    M   = round(mean(score, na.rm = TRUE), 2),
    SD  = round(sd(score, na.rm = TRUE),   2),
    Mdn = round(median(score, na.rm = TRUE), 2),
    Min = round(min(score, na.rm = TRUE),   2),
    Max = round(max(score, na.rm = TRUE),   2),
    Skew = round(psych::skew(score, na.rm = TRUE), 2),
    Kurt = round(psych::kurtosi(score, na.rm = TRUE), 2),
    .groups = "drop"
  ) |>
  dplyr::arrange(Need, Sample) |>
  flextable() |>
  flextable::merge_v(j = "Need") |>
  style_apa7() |>
  add_apa_note(paste0(
    "LLM score = mean of 3 scoring repeats. All scores on a 1–5 scale. ",
    "Skew = Fisher's skewness; Kurt = excess kurtosis."
  ))
```

## 7.2 LLM Score Distributions by Group

```{r}
#| label: fig-llm-dist-bygroup
#| echo: false
#| fig-cap: "LLM Score Distributions by Group and Need Dimension. Violin + boxplot. AI surrogates (blue) vs human participants (red)."
#| fig-width: 9
#| fig-height: 4
#| fig-dpi: 300

d_rq3_merged |>
  tidyr::pivot_longer(
    cols = c(llm_aut_mean, llm_com_mean, llm_rel_mean),
    names_to = "need_var", values_to = "score"
  ) |>
  dplyr::mutate(
    Need   = factor(dplyr::case_when(
      need_var == "llm_aut_mean" ~ TABLE_LABEL_AUT,
      need_var == "llm_com_mean" ~ TABLE_LABEL_COM,
      need_var == "llm_rel_mean" ~ TABLE_LABEL_REL
    ), levels = NEEDS),
    Sample = dplyr::recode(id_type, "AI" = LABEL_AI, "HU" = LABEL_HU)
  ) |>
  ggplot(aes(x = Sample, y = score, fill = id_type)) +
  geom_violin(alpha = 0.4, trim = FALSE, linewidth = 0.4) +
  geom_boxplot(width = 0.18, outlier.size = 0.8, outlier.alpha = 0.5) +
  scale_fill_manual(values = c("AI" = COL_AI_SINGLE, "HU" = COL_HU_SINGLE),
                    guide = "none") +
  scale_y_continuous(limits = c(1, 5), breaks = 1:5) +
  facet_wrap(~Need, ncol = 3L) +
  labs(x = NULL, y = "LLM score") +
  theme_thesis()
```

## 7.3 LLM vs Questionnaire Scatter by Group

```{r}
#| label: fig-llm-q-scatter-bygroup
#| echo: false
#| fig-cap: "LLM score vs questionnaire score by group and need dimension. LOESS (solid) and OLS (dashed) overlays. Supports assumption check for RQ3b."
#| fig-width: 9
#| fig-height: 6
#| fig-dpi: 300
#| warning: false

d_rq3_merged |>
  tidyr::pivot_longer(
    cols = c(llm_aut_mean, llm_com_mean, llm_rel_mean,
             q_aut_mean,   q_com_mean,   q_rel_mean),
    names_to  = "var", values_to = "score"
  ) |>
  dplyr::mutate(
    instrument = dplyr::if_else(stringr::str_starts(var, "llm_"), "llm", "q"),
    Need = factor(dplyr::case_when(
      stringr::str_detect(var, "aut") ~ TABLE_LABEL_AUT,
      stringr::str_detect(var, "com") ~ TABLE_LABEL_COM,
      stringr::str_detect(var, "rel") ~ TABLE_LABEL_REL
    ), levels = NEEDS),
    Sample = dplyr::recode(id_type, "AI" = LABEL_AI, "HU" = LABEL_HU)
  ) |>
  dplyr::select(-var) |>
  tidyr::pivot_wider(id_cols = c(id, id_type, Need, Sample),
                     names_from = instrument, values_from = score) |>
  ggplot(aes(x = llm, y = q, colour = id_type)) +
  geom_jitter(alpha = 0.25, size = 0.9, width = 0.12, height = 0.12) +
  geom_smooth(method = "loess", se = FALSE, linewidth = 0.7, formula = y ~ x) +
  geom_smooth(method = "lm",    se = FALSE, linewidth = 0.5, linetype = "dashed",
              formula = y ~ x) +
  scale_colour_manual(values = c("AI" = COL_AI_SINGLE, "HU" = COL_HU_SINGLE),
                      labels  = c("AI" = LABEL_AI, "HU" = LABEL_HU),
                      name    = NULL) +
  scale_x_continuous(limits = c(1, 5), breaks = 1:5) +
  scale_y_continuous(limits = c(1, 5), breaks = 1:5) +
  facet_grid(Need ~ Sample) +
  labs(x = "LLM score", y = "Questionnaire score") +
  theme_thesis() +
  theme(strip.text = element_text(size = 9))
```

---

# 8 Exports

```{r}
#| label: exports
#| include: false

# d_rq3_merged: merged AI + human (matched condition), minimal columns for RQ3 analyses
saveRDS(d_rq3_merged,  file.path(EXPORT_DIR, "d_rq3_merged.rds"))

# d_human_clean: full human dataset including demographics and item-level columns
saveRDS(d_human_clean, file.path(EXPORT_DIR, "d_human_clean.rds"))

message("Exports complete:")
message("  d_rq3_merged.rds  — ", nrow(d_rq3_merged), " rows (", N_AI, " AI + ", N_HU, " human)")
message("  d_human_clean.rds — ", nrow(d_human_clean), " rows")
```

---

# 9 Session Info

```{r}
#| label: session-info
#| include: false

sessionInfo()
```
