---
title: "RQ3 Analyses — Distributional Equivalence and Correlation Correspondence"
subtitle: "Bootstrap CIs · BCa sensitivity · Welch t · Levene · Fisher z · Spearman"
author: "Caroline Müller"
date: "2026-05-17"
format:
  html:
    code-fold: true
    code-summary: "Show code"
    code-tools: true
    toc: true
    self-contained: true
    fig-cap-location: top
editor:
  markdown:
    wrap: 72
knitr:
  opts_chunk:
    dpi: 300
---

# 1 Overview

**Prerequisite:** Run `30_RQ3_data.qmd` first to generate
`d_rq3_merged.rds` in `10_data/80_analyses_exports/`.

This file covers all inferential RQ3 analyses:

- **RQ3a** Distributional equivalence: delta-M (Welch t) and delta-SD
  (Levene's) with bootstrap percentile CIs
- **RQ3b** Correlation correspondence: Pearson r per group + Fisher's
  z-difference test + Spearman robustness
- **BCa sensitivity** checks for delta-SD and delta-r (bounded/skewed)

**Primary CIs:** Bootstrap percentile, R = 10,000, stratified by
`id_type`.\
**Sensitivity CIs:** BCa for delta-SD and delta-r; divergence threshold
= 0.02.

**Note:** The `id_type = "AI"` group is the short-3 matched condition
from `d_rq3_merged`. All inferential claims about means and SDs are for
this matched condition only; multi-condition AI comparisons are in
`20_RQ2_analyses.qmd`.

---

# 2 Setup

## 2.1 Environment

```{r}
#| label: setup-environment
#| include: false
#| warning: false

graphics.off()

suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(tidyr)
  library(flextable)
  library(scales)
  library(psych)
  library(officer)
  library(janitor)
  library(systemfonts)
  library(ragg)
  library(boot)
  library(car)
  library(purrr)
  library(cowplot)
  library(ggtext)
})

knitr::opts_chunk$set(dev = "ragg_png")
if (.Platform$OS.type == "windows") windowsFonts(Calibri = windowsFont("Calibri"))

message("Libraries loaded.")
```

## 2.2 Shared Constants

```{r}
#| label: setup-common
#| include: false

source("00_common.R")
```

## 2.3 Load Upstream Export

```{r}
#| label: setup-imports
#| include: false

EXPORT_DIR <- "../10_data/80_analyses_exports"

rds_paths <- list(
  merged = file.path(EXPORT_DIR, "d_rq3_merged.rds")
)

for (nm in names(rds_paths)) {
  if (!file.exists(rds_paths[[nm]])) {
    stop("Missing upstream export — run 30_RQ3_data.qmd first: ", rds_paths[[nm]])
  }
}

d_rq3_merged <- readRDS(rds_paths$merged)

N_AI <- sum(d_rq3_merged$id_type == "AI")
N_HU <- sum(d_rq3_merged$id_type == "HU")

message("d_rq3_merged loaded: ", N_AI, " AI + ", N_HU, " human = ",
        nrow(d_rq3_merged), " total rows.")
```

## 2.4 Helpers

```{r}
#| label: setup-helpers
#| include: false

# Point estimate + percentile CI from a numeric vector of bootstrap replicates
perc_ci <- function(boots, probs = c(CI_Q_LO, CI_Q_HI)) {
  quantile(boots, probs = probs, na.rm = TRUE)
}

message("Helpers defined.")
```

---

# 3 Assumption Checks

## 3.1 Univariate Normality

Shapiro-Wilk at N ≈ 400 is overpowered; significant results are expected
for trivially small deviations. We use Q-Q plots as the primary
diagnostic; skewness and kurtosis are reported for reference.

```{r}
#| label: fig-assump-qq
#| echo: false
#| fig-cap: "Univariate Q-Q Plots by Group and Need Dimension. Substantial departures from the diagonal support non-normality."
#| fig-width: 9
#| fig-height: 6
#| fig-dpi: 300

long_rq3 <- d_rq3_merged |>
  tidyr::pivot_longer(
    cols      = c(llm_aut_mean, llm_com_mean, llm_rel_mean,
                  q_aut_mean,   q_com_mean,   q_rel_mean),
    names_to  = "variable",
    values_to = "score"
  ) |>
  dplyr::mutate(
    Instrument = dplyr::if_else(startsWith(variable, "llm_"), "LLM score", "Q score"),
    Need = factor(dplyr::case_when(
      grepl("aut", variable) ~ TABLE_LABEL_AUT,
      grepl("com", variable) ~ TABLE_LABEL_COM,
      grepl("rel", variable) ~ TABLE_LABEL_REL
    ), levels = NEEDS),
    Sample = dplyr::recode(id_type, "AI" = LABEL_AI, "HU" = LABEL_HU)
  )

long_rq3 |>
  dplyr::filter(Instrument == "LLM score") |>
  ggplot(aes(sample = score, colour = Sample)) +
  stat_qq(size = 0.7, alpha = 0.5) +
  stat_qq_line(linewidth = 0.5) +
  scale_colour_manual(values = c(setNames(COL_AI_SINGLE, LABEL_AI),
                                 setNames(COL_HU_SINGLE, LABEL_HU)),
                      name = NULL) +
  facet_grid(Sample ~ Need) +
  labs(x = "Theoretical quantiles", y = "Sample quantiles") +
  theme_thesis()
```

## 3.2 Linearity Check

```{r}
#| label: fig-assump-scatter-rq3
#| echo: false
#| fig-cap: "Linearity Check: LLM score vs questionnaire score by group and need. LOESS (solid) vs OLS (dashed)."
#| fig-width: 9
#| fig-height: 6
#| fig-dpi: 300

d_rq3_merged |>
  tidyr::pivot_longer(
    cols = c(llm_aut_mean, llm_com_mean, llm_rel_mean,
             q_aut_mean,   q_com_mean,   q_rel_mean),
    names_to  = "var", values_to = "score"
  ) |>
  dplyr::mutate(
    instrument = dplyr::if_else(startsWith(var, "llm_"), "llm", "q"),
    Need = factor(dplyr::case_when(
      grepl("aut", var) ~ TABLE_LABEL_AUT,
      grepl("com", var) ~ TABLE_LABEL_COM,
      grepl("rel", var) ~ TABLE_LABEL_REL
    ), levels = NEEDS),
    Sample = dplyr::recode(id_type, "AI" = LABEL_AI, "HU" = LABEL_HU)
  ) |>
  dplyr::select(-var) |>
  tidyr::pivot_wider(id_cols = c(id, id_type, Need, Sample),
                     names_from = instrument, values_from = score) |>
  ggplot(aes(x = llm, y = q, colour = id_type)) +
  geom_jitter(alpha = 0.25, size = 0.9, width = 0.12, height = 0.12) +
  geom_smooth(method = "loess", se = FALSE, linewidth = 0.7, formula = y ~ x) +
  geom_smooth(method = "lm",    se = FALSE, linewidth = 0.5, linetype = "dashed",
              formula = y ~ x) +
  scale_colour_manual(values = c("AI" = COL_AI_SINGLE, "HU" = COL_HU_SINGLE),
                      labels  = c("AI" = LABEL_AI, "HU" = LABEL_HU), name = NULL) +
  scale_x_continuous(limits = c(1, 5), breaks = 1:5) +
  scale_y_continuous(limits = c(1, 5), breaks = 1:5) +
  facet_grid(Need ~ Sample) +
  labs(x = "LLM score", y = "Questionnaire score") +
  theme_thesis()
```

---

# 4 RQ3a — Distributional Equivalence

RQ3a asks: to what extent do distributional properties (mean, SD) differ
between AI surrogate and human participant data?

## 4.1 Bootstrap delta-M

Stratified bootstrap: resample within AI and human groups separately.
Set.seed(2026) before each `boot::boot()` call.

```{r}
#| label: boot-delta-m-rq3
#| include: false
#| cache: true
#| cache.extra: !expr list(file.mtime(rds_paths$merged), BOOTSTRAP_SAMPLES)

boot_fn_delta_m <- function(data, indices) {
  d <- data[indices, ]
  ai <- dplyr::filter(d, id_type == "AI")
  hu <- dplyr::filter(d, id_type == "HU")
  c(
    aut = mean(ai$llm_aut_mean, na.rm = TRUE) - mean(hu$llm_aut_mean, na.rm = TRUE),
    com = mean(ai$llm_com_mean, na.rm = TRUE) - mean(hu$llm_com_mean, na.rm = TRUE),
    rel = mean(ai$llm_rel_mean, na.rm = TRUE) - mean(hu$llm_rel_mean, na.rm = TRUE)
  )
}

set.seed(2026)
boot_delta_m <- boot::boot(
  data      = d_rq3_merged,
  statistic = boot_fn_delta_m,
  R         = BOOTSTRAP_SAMPLES,
  strata    = as.integer(factor(d_rq3_merged$id_type))
)
colnames(boot_delta_m$t) <- names(boot_delta_m$t0)

saveRDS(boot_delta_m, file.path(EXPORT_DIR, "boot_delta_m.rds"))
message("boot_delta_m complete and saved.")
```

## 4.2 BCa Sensitivity Check — delta-M

```{r}
#| label: check-bca-delta-m
#| include: false
#| dependson: boot-delta-m-rq3

bca_delta_m_results <- purrr::map_dfr(
  list(list(idx = 1L, label = TABLE_LABEL_AUT),
       list(idx = 2L, label = TABLE_LABEL_COM),
       list(idx = 3L, label = TABLE_LABEL_REL)),
  function(x) {
    perc_ci_val <- boot::boot.ci(boot_delta_m, index = x$idx, type = "perc")$percent[4:5]
    bca_ci_val  <- tryCatch(
      boot::boot.ci(boot_delta_m, index = x$idx, type = "bca")$bca[4:5],
      error = function(e) c(NA_real_, NA_real_)
    )
    tibble::tibble(
      need    = x$label,
      perc_lo = perc_ci_val[1], perc_hi = perc_ci_val[2],
      bca_lo  = bca_ci_val[1],  bca_hi  = bca_ci_val[2],
      max_diff = max(abs(bca_lo - perc_lo), abs(bca_hi - perc_hi), na.rm = TRUE)
    )
  }
)

check_bca_divergence_delta_m <- bca_delta_m_results |> dplyr::filter(max_diff > 0.02)

if (nrow(check_bca_divergence_delta_m) > 0) {
  cat("BCa/percentile divergence > .02 in", nrow(check_bca_divergence_delta_m),
      "delta-M cell(s):\n")
  print(check_bca_divergence_delta_m)
} else {
  cat("BCa sensitivity check OK: all delta-M CIs agree within .02.\n")
}
```

## 4.3 Bootstrap delta-SD

```{r}
#| label: boot-delta-sd-rq3
#| include: false
#| cache: true
#| cache.extra: !expr list(file.mtime(rds_paths$merged), BOOTSTRAP_SAMPLES)

boot_fn_delta_sd <- function(data, indices) {
  d <- data[indices, ]
  ai <- dplyr::filter(d, id_type == "AI")
  hu <- dplyr::filter(d, id_type == "HU")
  c(
    aut = sd(ai$llm_aut_mean, na.rm = TRUE) - sd(hu$llm_aut_mean, na.rm = TRUE),
    com = sd(ai$llm_com_mean, na.rm = TRUE) - sd(hu$llm_com_mean, na.rm = TRUE),
    rel = sd(ai$llm_rel_mean, na.rm = TRUE) - sd(hu$llm_rel_mean, na.rm = TRUE)
  )
}

set.seed(2026)
boot_delta_sd <- boot::boot(
  data      = d_rq3_merged,
  statistic = boot_fn_delta_sd,
  R         = BOOTSTRAP_SAMPLES,
  strata    = as.integer(factor(d_rq3_merged$id_type))
)
colnames(boot_delta_sd$t) <- names(boot_delta_sd$t0)

saveRDS(boot_delta_sd, file.path(EXPORT_DIR, "boot_delta_sd.rds"))
message("boot_delta_sd complete and saved.")
```

## 4.4 BCa Sensitivity Check — delta-SD

```{r}
#| label: check-bca-delta-sd
#| include: false
#| dependson: boot-delta-sd-rq3

bca_delta_sd_results <- purrr::map_dfr(
  list(list(idx = 1L, label = TABLE_LABEL_AUT),
       list(idx = 2L, label = TABLE_LABEL_COM),
       list(idx = 3L, label = TABLE_LABEL_REL)),
  function(x) {
    perc_ci_val <- boot::boot.ci(boot_delta_sd, index = x$idx, type = "perc")$percent[4:5]
    bca_ci_val  <- tryCatch(
      boot::boot.ci(boot_delta_sd, index = x$idx, type = "bca")$bca[4:5],
      error = function(e) c(NA_real_, NA_real_)
    )
    tibble::tibble(
      need    = x$label,
      perc_lo = perc_ci_val[1], perc_hi = perc_ci_val[2],
      bca_lo  = bca_ci_val[1],  bca_hi  = bca_ci_val[2],
      max_diff = max(abs(bca_lo - perc_lo), abs(bca_hi - perc_hi), na.rm = TRUE)
    )
  }
)

check_bca_divergence_delta_sd <- bca_delta_sd_results |> dplyr::filter(max_diff > 0.02)

if (nrow(check_bca_divergence_delta_sd) > 0) {
  cat("BCa/percentile divergence > .02 in", nrow(check_bca_divergence_delta_sd),
      "delta-SD cell(s):\n")
  print(check_bca_divergence_delta_sd)
} else {
  cat("BCa sensitivity check OK: all delta-SD CIs agree within .02.\n")
}
```

## 4.5 Welch t-test and Levene's Test

```{r}
#| label: compute-rq3a-tests
#| include: false

# Welch t-test (delta-M) and Levene's test (delta-SD, center="mean") per need
rq3a_tests <- purrr::map_dfr(
  list(list(llm = "llm_aut_mean", need = TABLE_LABEL_AUT, idx = 1L),
       list(llm = "llm_com_mean", need = TABLE_LABEL_COM, idx = 2L),
       list(llm = "llm_rel_mean", need = TABLE_LABEL_REL, idx = 3L)),
  function(x) {
    ai_vals <- d_rq3_merged[d_rq3_merged$id_type == "AI", x$llm][[1]]
    hu_vals <- d_rq3_merged[d_rq3_merged$id_type == "HU", x$llm][[1]]

    # Welch t-test (unequal variances assumed)
    t_res   <- t.test(ai_vals, hu_vals, var.equal = FALSE)

    # Levene's test on the full grouped data (center = "mean" per pre-registration)
    lev_res <- car::leveneTest(
      d_rq3_merged[[x$llm]] ~ factor(d_rq3_merged$id_type),
      center = "mean"
    )

    # Bootstrap percentile CIs for delta-M
    delta_m_perc <- perc_ci(boot_delta_m$t[, x$idx])
    delta_sd_perc <- perc_ci(boot_delta_sd$t[, x$idx])

    tibble::tibble(
      Need         = x$need,
      n_AI         = sum(!is.na(ai_vals)),
      n_HU         = sum(!is.na(hu_vals)),
      M_AI         = round(mean(ai_vals, na.rm = TRUE), 2),
      M_HU         = round(mean(hu_vals, na.rm = TRUE), 2),
      SD_AI        = round(sd(ai_vals, na.rm = TRUE), 2),
      SD_HU        = round(sd(hu_vals, na.rm = TRUE), 2),
      delta_M      = sprintf("%.2f [%.2f, %.2f]",
                             t_res$estimate[1] - t_res$estimate[2],
                             delta_m_perc[1], delta_m_perc[2]),
      t_val        = round(t_res$statistic, 2),
      df_t         = round(t_res$parameter, 1),
      p_t          = round(t_res$p.value, 3),
      delta_SD     = sprintf("%.2f [%.2f, %.2f]",
                             sd(ai_vals, na.rm = TRUE) - sd(hu_vals, na.rm = TRUE),
                             delta_sd_perc[1], delta_sd_perc[2]),
      Levene_F     = round(lev_res$`F value`[1], 2),
      df_lev       = paste0(lev_res$Df[1], ", ", lev_res$Df[2]),
      p_lev        = round(lev_res$`Pr(>F)`[1], 3)
    )
  }
)

message("RQ3a tests computed.")
```

## 4.6 RQ3a Results Table

```{r}
#| label: tbl-rq3a-results
#| echo: false
#| tbl-cap: "RQ3a: Distributional Equivalence — Welch t-test and Levene's Test by Need Dimension"

rq3a_tests |>
  dplyr::select(Need, n_AI, n_HU, M_AI, M_HU, SD_AI, SD_HU,
                delta_M, t_val, df_t, p_t,
                delta_SD, Levene_F, df_lev, p_lev) |>
  flextable() |>
  flextable::set_header_labels(
    n_AI       = "N (AI)",  n_HU      = "N (HU)",
    M_AI       = "M (AI)",  M_HU      = "M (HU)",
    SD_AI      = "SD (AI)", SD_HU     = "SD (HU)",
    delta_M    = "Δ M [95% CI]",
    t_val      = "t",       df_t      = "df",  p_t     = "p",
    delta_SD   = "Δ SD [95% CI]",
    Levene_F   = "F",       df_lev    = "df",  p_lev   = "p"
  ) |>
  style_apa7() |>
  add_apa_note(paste0(
    "Δ M = M(AI) − M(HU); Δ SD = SD(AI) − SD(HU); descriptive effect magnitudes, not inputs to the inferential tests. ",
    "Bootstrap percentile CIs, R = ", BOOTSTRAP_SAMPLES, ", stratified by group, seed = 2026. ",
    "Welch's t-test uses group means, SDs, and Ns via the Welch–Satterthwaite formula (var.equal = FALSE). ",
    "Levene's test uses mean-absolute deviations from group means (center = mean). ",
    "BCa sensitivity: see check_bca_divergence_delta_m, check_bca_divergence_delta_sd."
  ))
```

## 4.7 RQ3a Bar Plot

```{r}
#| label: fig-rq3a-bars
#| echo: false
#| fig-cap: "RQ3a: LLM Score Means by Group and Need Dimension. Error bars = approximate 95% CI (M ± 1.96 SE, normal theory, illustrative). Inferential tests in tbl-rq3a-results."
#| fig-width: 7
#| fig-height: 4
#| fig-dpi: 300

# Use rq3a_tests for point estimates; error bars = ±1.96 SE (illustrative only)
rq3a_plot_data <- rq3a_tests |>
  tidyr::pivot_longer(
    cols      = c(M_AI, M_HU, SD_AI, SD_HU, n_AI, n_HU),
    names_to  = c(".value", "id_type"),
    names_sep = "_"
  ) |>
  dplyr::mutate(
    Need  = factor(Need, levels = NEEDS),
    ci_lo = M - CI_BOUND_95 * SD / sqrt(n),
    ci_hi = M + CI_BOUND_95 * SD / sqrt(n)
  )

ggplot(rq3a_plot_data,
       aes(x = Need, y = M, fill = id_type, group = id_type)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_errorbar(aes(ymin = ci_lo, ymax = ci_hi),
                position = position_dodge(width = 0.7),
                width = 0.2, colour = COLOR_ERRORBAR) +
  scale_fill_manual(values = c("AI" = COL_AI_SINGLE, "HU" = COL_HU_SINGLE),
                    labels  = c("AI" = LABEL_AI, "HU" = LABEL_HU),
                    name    = NULL) +
  scale_y_continuous(limits = c(0, 5), breaks = 0:5) +
  labs(x = LABEL_NEEDS, y = "Mean LLM score") +
  theme_thesis()
```

## 4.8 RQ3a Distribution Plot

```{r}
#| label: fig-rq3a-distributions
#| echo: false
#| fig-cap: "RQ3a: LLM Score Distributions by Group and Need Dimension. Overlapping density curves for AI surrogates and human participants."
#| fig-width: 9
#| fig-height: 4
#| fig-dpi: 300

d_rq3_merged |>
  tidyr::pivot_longer(
    cols      = c(llm_aut_mean, llm_com_mean, llm_rel_mean),
    names_to  = "variable",
    values_to = "score"
  ) |>
  dplyr::mutate(
    Need = factor(dplyr::case_when(
      grepl("aut", variable) ~ TABLE_LABEL_AUT,
      grepl("com", variable) ~ TABLE_LABEL_COM,
      grepl("rel", variable) ~ TABLE_LABEL_REL
    ), levels = NEEDS)
  ) |>
  ggplot(aes(x = score, fill = id_type, colour = id_type)) +
  geom_density(alpha = 0.30, linewidth = 0.6) +
  scale_fill_manual(values = c("AI" = COL_AI_SINGLE, "HU" = COL_HU_SINGLE),
                    labels  = c("AI" = LABEL_AI, "HU" = LABEL_HU), name = NULL) +
  scale_colour_manual(values = c("AI" = COL_AI_SINGLE, "HU" = COL_HU_SINGLE),
                      labels  = c("AI" = LABEL_AI, "HU" = LABEL_HU), name = NULL) +
  scale_x_continuous(limits = c(1, 5), breaks = 1:5) +
  facet_wrap(~Need, ncol = 3L) +
  labs(x = "Mean LLM score", y = "Density") +
  theme_thesis()
```

## 4.9 RQ3a Delta-M and Delta-SD Forest Plots

```{r}
#| label: fig-rq3a-delta-forest
#| echo: false
#| fig-cap: "RQ3a: Bootstrap Estimates of Δ M (left) and Δ SD (right) by Need Dimension. Points = observed AI − Human difference; horizontal segments = 95% bootstrap percentile CIs; dashed line = 0."
#| fig-width: 9
#| fig-height: 3
#| fig-dpi: 300

delta_m_forest_data <- tibble::tibble(
  Need     = factor(NEEDS, levels = rev(NEEDS)),
  estimate = as.numeric(boot_delta_m$t0),
  ci_lo    = apply(boot_delta_m$t, 2L, quantile, CI_Q_LO, na.rm = TRUE),
  ci_hi    = apply(boot_delta_m$t, 2L, quantile, CI_Q_HI, na.rm = TRUE)
)

delta_sd_forest_data <- tibble::tibble(
  Need     = factor(NEEDS, levels = rev(NEEDS)),
  estimate = as.numeric(boot_delta_sd$t0),
  ci_lo    = apply(boot_delta_sd$t, 2L, quantile, CI_Q_LO, na.rm = TRUE),
  ci_hi    = apply(boot_delta_sd$t, 2L, quantile, CI_Q_HI, na.rm = TRUE)
)

forest_theme_rq3 <- theme_thesis() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(linewidth = 0.2)
  )

p_dm <- ggplot(delta_m_forest_data, aes(x = estimate, y = Need)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  geom_segment(aes(x = ci_lo, xend = ci_hi, y = Need, yend = Need), linewidth = 0.8) +
  geom_point(size = 3, shape = 21, fill = "white", stroke = 1.8) +
  labs(x = "Δ M (AI − Human) [95% CI]", y = NULL) +
  forest_theme_rq3

p_dsd <- ggplot(delta_sd_forest_data, aes(x = estimate, y = Need)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  geom_segment(aes(x = ci_lo, xend = ci_hi, y = Need, yend = Need), linewidth = 0.8) +
  geom_point(size = 3, shape = 21, fill = "white", stroke = 1.8) +
  labs(x = "Δ SD (AI − Human) [95% CI]", y = NULL) +
  forest_theme_rq3

cowplot::plot_grid(p_dm, p_dsd, ncol = 2L, align = "hv")
```

---

# 5 RQ3b — Correlation Correspondence

RQ3b asks: to what extent do LLM–questionnaire correlations differ
between AI surrogates and human participants?

## 5.1 Bootstrap delta-r

```{r}
#| label: boot-delta-r-rq3
#| include: false
#| cache: true
#| cache.extra: !expr list(file.mtime(rds_paths$merged), BOOTSTRAP_SAMPLES)

boot_fn_delta_r <- function(data, indices) {
  d  <- data[indices, ]
  ai <- dplyr::filter(d, id_type == "AI")
  hu <- dplyr::filter(d, id_type == "HU")
  r_ai_aut <- cor(ai$llm_aut_mean, ai$q_aut_mean, use = "complete.obs")
  r_hu_aut <- cor(hu$llm_aut_mean, hu$q_aut_mean, use = "complete.obs")
  r_ai_com <- cor(ai$llm_com_mean, ai$q_com_mean, use = "complete.obs")
  r_hu_com <- cor(hu$llm_com_mean, hu$q_com_mean, use = "complete.obs")
  r_ai_rel <- cor(ai$llm_rel_mean, ai$q_rel_mean, use = "complete.obs")
  r_hu_rel <- cor(hu$llm_rel_mean, hu$q_rel_mean, use = "complete.obs")
  c(
    aut = r_ai_aut - r_hu_aut,
    com = r_ai_com - r_hu_com,
    rel = r_ai_rel - r_hu_rel
  )
}

set.seed(2026)
boot_delta_r <- boot::boot(
  data      = d_rq3_merged,
  statistic = boot_fn_delta_r,
  R         = BOOTSTRAP_SAMPLES,
  strata    = as.integer(factor(d_rq3_merged$id_type))
)
colnames(boot_delta_r$t) <- names(boot_delta_r$t0)

saveRDS(boot_delta_r, file.path(EXPORT_DIR, "boot_delta_r.rds"))
message("boot_delta_r complete and saved.")
```

## 5.2 BCa Sensitivity Check — delta-r

```{r}
#| label: check-bca-delta-r
#| include: false
#| dependson: boot-delta-r-rq3

bca_delta_r_results <- purrr::map_dfr(
  list(list(idx = 1L, label = TABLE_LABEL_AUT),
       list(idx = 2L, label = TABLE_LABEL_COM),
       list(idx = 3L, label = TABLE_LABEL_REL)),
  function(x) {
    perc_ci_val <- boot::boot.ci(boot_delta_r, index = x$idx, type = "perc")$percent[4:5]
    bca_ci_val  <- tryCatch(
      boot::boot.ci(boot_delta_r, index = x$idx, type = "bca")$bca[4:5],
      error = function(e) c(NA_real_, NA_real_)
    )
    tibble::tibble(
      need    = x$label,
      perc_lo = perc_ci_val[1], perc_hi = perc_ci_val[2],
      bca_lo  = bca_ci_val[1],  bca_hi  = bca_ci_val[2],
      max_diff = max(abs(bca_lo - perc_lo), abs(bca_hi - perc_hi), na.rm = TRUE)
    )
  }
)

check_bca_divergence_delta_r <- bca_delta_r_results |> dplyr::filter(max_diff > 0.02)

if (nrow(check_bca_divergence_delta_r) > 0) {
  cat("BCa/percentile divergence > .02 in", nrow(check_bca_divergence_delta_r),
      "delta-r cell(s):\n")
  print(check_bca_divergence_delta_r)
} else {
  cat("BCa sensitivity check OK: all delta-r CIs agree within .02.\n")
}
```

## 5.3 Fisher z-difference Test and Results

```{r}
#| label: compute-rq3b-tests
#| include: false

rq3b_tests <- purrr::map_dfr(
  list(list(llm = "llm_aut_mean", q = "q_aut_mean", need = TABLE_LABEL_AUT, idx = 1L),
       list(llm = "llm_com_mean", q = "q_com_mean", need = TABLE_LABEL_COM, idx = 2L),
       list(llm = "llm_rel_mean", q = "q_rel_mean", need = TABLE_LABEL_REL, idx = 3L)),
  function(x) {
    ai_dat <- d_rq3_merged[d_rq3_merged$id_type == "AI", ]
    hu_dat <- d_rq3_merged[d_rq3_merged$id_type == "HU", ]

    r_ai <- cor(ai_dat[[x$llm]], ai_dat[[x$q]], use = "complete.obs")
    r_hu <- cor(hu_dat[[x$llm]], hu_dat[[x$q]], use = "complete.obs")
    n_ai <- sum(complete.cases(ai_dat[[x$llm]], ai_dat[[x$q]]))
    n_hu <- sum(complete.cases(hu_dat[[x$llm]], hu_dat[[x$q]]))

    # Fisher z-difference test: psych::r.test()
    rtest <- psych::r.test(n = n_ai, n2 = n_hu, r12 = r_ai, r34 = r_hu)

    delta_r_perc <- perc_ci(boot_delta_r$t[, x$idx])

    tibble::tibble(
      Need        = x$need,
      r_AI        = round(r_ai, 2),
      r_HU        = round(r_hu, 2),
      n_AI        = n_ai,
      n_HU        = n_hu,
      delta_r     = round(r_ai - r_hu, 2),
      delta_r_ci  = sprintf("[%.2f, %.2f]", delta_r_perc[1], delta_r_perc[2]),
      Fisher_z    = round(rtest$z, 2),
      p_Fisher    = round(rtest$p, 3)
    )
  }
)

message("RQ3b tests computed.")
```

## 5.4 RQ3b Results Table

```{r}
#| label: tbl-rq3b-results
#| echo: false
#| tbl-cap: "RQ3b: Correlation Correspondence — Pearson r per Group and Fisher z-Difference by Need Dimension"

rq3b_tests |>
  flextable() |>
  flextable::set_header_labels(
    r_AI       = "r (AI)", r_HU       = "r (HU)",
    n_AI       = "N (AI)", n_HU       = "N (HU)",
    delta_r    = "Δ r",    delta_r_ci  = "95% CI (Δ r)",
    Fisher_z   = "z",      p_Fisher    = "p"
  ) |>
  style_apa7() |>
  add_apa_note(paste0(
    "r = Pearson LLM–questionnaire correlation. ",
    "Δ r = r(AI) − r(HU); bootstrap percentile CIs, R = ", BOOTSTRAP_SAMPLES,
    ", stratified by group, seed = 2026. ",
    "Fisher z-difference test via psych::r.test(), two independent groups. ",
    "BCa sensitivity: see check_bca_divergence_delta_r."
  ))
```

## 5.5 RQ3b Scatter Plot

```{r}
#| label: fig-rq3b-scatter
#| echo: false
#| fig-cap: "RQ3b: LLM score vs questionnaire score by group and need dimension. LOESS (solid), OLS (dashed). Pearson r annotated per panel."
#| fig-width: 9
#| fig-height: 6
#| fig-dpi: 300

# Annotate each panel with r value
r_labels <- dplyr::bind_rows(
  rq3b_tests |>
    dplyr::transmute(
      id_type = "AI", Need = factor(Need, levels = NEEDS),
      label = sprintf("r = %.2f", r_AI),
      x = 1.3, y = 4.7
    ),
  rq3b_tests |>
    dplyr::transmute(
      id_type = "HU", Need = factor(Need, levels = NEEDS),
      label = sprintf("r = %.2f", r_HU),
      x = 1.3, y = 4.7
    )
)

d_rq3_merged |>
  tidyr::pivot_longer(
    cols = c(llm_aut_mean, llm_com_mean, llm_rel_mean,
             q_aut_mean,   q_com_mean,   q_rel_mean),
    names_to  = "var", values_to = "score"
  ) |>
  dplyr::mutate(
    instrument = dplyr::if_else(startsWith(var, "llm_"), "llm", "q"),
    Need = factor(dplyr::case_when(
      grepl("aut", var) ~ TABLE_LABEL_AUT,
      grepl("com", var) ~ TABLE_LABEL_COM,
      grepl("rel", var) ~ TABLE_LABEL_REL
    ), levels = NEEDS)
  ) |>
  dplyr::select(-var) |>
  tidyr::pivot_wider(id_cols = c(id, id_type, Need),
                     names_from = instrument, values_from = score) |>
  ggplot(aes(x = llm, y = q, colour = id_type)) +
  geom_jitter(alpha = 0.25, size = 0.9, width = 0.12, height = 0.12) +
  geom_smooth(method = "loess", se = FALSE, linewidth = 0.7, formula = y ~ x) +
  geom_smooth(method = "lm",    se = FALSE, linewidth = 0.5, linetype = "dashed",
              formula = y ~ x) +
  geom_text(data = r_labels, aes(x = x, y = y, label = label, colour = id_type),
            size = 3, hjust = 0, family = THESIS_FONT, inherit.aes = FALSE) +
  scale_colour_manual(values = c("AI" = COL_AI_SINGLE, "HU" = COL_HU_SINGLE),
                      labels  = c("AI" = LABEL_AI, "HU" = LABEL_HU), name = NULL) +
  scale_x_continuous(limits = c(1, 5), breaks = 1:5) +
  scale_y_continuous(limits = c(1, 5), breaks = 1:5) +
  facet_grid(Need ~ id_type,
             labeller = labeller(id_type = c("AI" = LABEL_AI, "HU" = LABEL_HU))) +
  labs(x = "LLM score", y = "Questionnaire score") +
  theme_thesis() +
  theme(legend.position = "none")
```

## 5.5.1 Bootstrap Per-Group r CIs

```{r}
#| label: boot-rq3b-group-cis
#| include: false

boot_fn_r_single <- function(data, i) cor(data$llm[i], data$q[i], method = "pearson")

rq3b_group_ci_data <- purrr::map_dfr(
  list(list(llm = "llm_aut_mean", q = "q_aut_mean", need = TABLE_LABEL_AUT),
       list(llm = "llm_com_mean", q = "q_com_mean", need = TABLE_LABEL_COM),
       list(llm = "llm_rel_mean", q = "q_rel_mean", need = TABLE_LABEL_REL)),
  function(x) {
    ai_dat <- d_rq3_merged[d_rq3_merged$id_type == "AI", ]
    hu_dat <- d_rq3_merged[d_rq3_merged$id_type == "HU", ]

    r_ai <- cor(ai_dat[[x$llm]], ai_dat[[x$q]], use = "complete.obs")
    r_hu <- cor(hu_dat[[x$llm]], hu_dat[[x$q]], use = "complete.obs")

    set.seed(2026)
    boot_ai <- boot::boot(
      data.frame(llm = ai_dat[[x$llm]], q = ai_dat[[x$q]]),
      boot_fn_r_single, R = BOOTSTRAP_SAMPLES)
    ci_ai <- boot::boot.ci(boot_ai, type = "perc")$percent[4:5]

    set.seed(2026)
    boot_hu <- boot::boot(
      data.frame(llm = hu_dat[[x$llm]], q = hu_dat[[x$q]]),
      boot_fn_r_single, R = BOOTSTRAP_SAMPLES)
    ci_hu <- boot::boot.ci(boot_hu, type = "perc")$percent[4:5]

    dplyr::bind_rows(
      tibble::tibble(need = x$need, group = LABEL_AI, r = r_ai,
                     ci_lo = ci_ai[1], ci_hi = ci_ai[2]),
      tibble::tibble(need = x$need, group = LABEL_HU, r = r_hu,
                     ci_lo = ci_hu[1], ci_hi = ci_hu[2])
    )
  }
) |>
  dplyr::mutate(
    need  = factor(need,  levels = c(TABLE_LABEL_AUT, TABLE_LABEL_COM, TABLE_LABEL_REL)),
    group = factor(group, levels = c(LABEL_HU, LABEL_AI))
  )

message("Per-group r bootstrap CIs computed.")
```

## 5.5.2 Figure: Correlations per Group — Dot Plot with CI

```{r}
#| label: fig-rq3b-corr-CI
#| echo: false
#| fig-cap: "Pearson r between LLM scores and questionnaire scores by need dimension and group (RQ3b). Error bars = 95% bootstrap CIs. Bracket = Δr (r_AI − r_Human)."
#| fig-width: 6
#| fig-height: 4
#| fig-dpi: 300

DODGE_W <- 0.1
TICK_W  <- 0.10

fig_rq3b_corr_ci_data <- rq3b_group_ci_data |>
  dplyr::mutate(x = as.numeric(need) + ifelse(group == LABEL_HU, -DODGE_W, DODGE_W))

bracket_data <- rq3b_group_ci_data |>
  dplyr::select(need, group, r) |>
  tidyr::pivot_wider(names_from = group, values_from = r) |>
  dplyr::rename(Need = need, r_AI = !!LABEL_AI, r_HU = !!LABEL_HU) |>
  dplyr::mutate(
    need_num = dplyr::case_when(
      Need == TABLE_LABEL_AUT ~ 1L,
      Need == TABLE_LABEL_COM ~ 2L,
      Need == TABLE_LABEL_REL ~ 3L
    ),
    x_arm = need_num
  )

rq3b_group_colors_ci <- c(
  setNames(COL_AI_SHADES2[4], LABEL_AI),
  setNames(COL_HU_SHADES[1], LABEL_HU)
  #COL_AI_SHADES2
)

ggplot(fig_rq3b_corr_ci_data,
    aes(x = x, y = r, color = group, fill = group, group = group)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.5, color = "grey60") +
  geom_errorbar(aes(ymin = ci_lo, ymax = ci_hi), width = 0, linewidth = 0.8) +
  geom_point(shape = 21, fill = "white", size = 1.5, stroke = 1.8) +
  geom_segment(data = bracket_data,
    aes(x = x_arm, xend = x_arm, y = r_HU, yend = r_AI),
    color = COL_DIFF, linewidth = 0.25, inherit.aes = FALSE) +
  geom_segment(data = bracket_data,
    aes(x = x_arm - TICK_W, xend = x_arm + TICK_W, y = r_AI, yend = r_AI),
    color = COL_DIFF, linewidth = 0.25, inherit.aes = FALSE) +
  geom_segment(data = bracket_data,
    aes(x = x_arm - TICK_W, xend = x_arm + TICK_W, y = r_HU, yend = r_HU),
    color = COL_DIFF, linewidth = 0.25, inherit.aes = FALSE) +
  scale_x_continuous(
    breaks   = 1:3,
    labels   = levels(fig_rq3b_corr_ci_data$need),
    limits   = c(0.6, 3.6),
    position = "top"
  ) +
  scale_color_manual(values = rq3b_group_colors_ci, name = NULL) +
  scale_fill_manual(values  = rq3b_group_colors_ci, name = NULL) +
  labs(x = NULL, y = "Pearson *r*") +
  theme_thesis() +
  theme(
    panel.grid.major.x   = element_blank(),
    panel.grid.minor.x   = element_blank(),
    axis.text.x.top      = element_text(colour = "black"),
    panel.grid.major.y   = element_line(linewidth = 0.3),
    panel.grid.minor.y   = element_line(linewidth = 0.3),
    axis.title.y         = ggtext::element_markdown(family = THESIS_FONT, margin = margin(r = 12)),
    legend.position      = c(0.18, 0.15),
    legend.direction     = "horizontal",
    legend.margin        = margin(t = 4),
    legend.background    = element_rect(fill = "white", color = NA),
    legend.key           = element_blank()
  )
```

*Note.* Error bars = 95% bootstrap CIs (R = `r BOOTSTRAP_SAMPLES`). AI surrogates N = `r N_AI`; Humans N = `r N_HU`.

```{r}
#| label: fig-rq3b-corr-CI-rect
#| echo: false
#| fig-cap: "Pearson r between LLM scores and questionnaire scores by need dimension and group (RQ3b). Error bars = 95% bootstrap CIs. Shaded rectangles = difference between point estimates (Δr)."
#| fig-width: 6
#| fig-height: 4
#| fig-dpi: 300

RECT_HALF_W <- 0.04
LABEL_DIFF  <- "Δr"

bracket_data_rect <- bracket_data |>
  dplyr::mutate(diff_label = LABEL_DIFF)

all_levels       <- c(LABEL_HU, LABEL_AI, LABEL_DIFF)
all_color_values <- c(rq3b_group_colors_ci, setNames("transparent", LABEL_DIFF))
all_fill_values  <- c(rq3b_group_colors_ci, setNames(COL_DIFF,      LABEL_DIFF))

ggplot(fig_rq3b_corr_ci_data,
    aes(x = x, y = r, color = group, fill = group, group = group)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.5, color = "grey60") +
  geom_rect(data = bracket_data_rect,
    aes(xmin = x_arm - RECT_HALF_W, xmax = x_arm + RECT_HALF_W,
        ymin = pmin(r_HU, r_AI),     ymax = pmax(r_HU, r_AI),
        fill = diff_label, color = diff_label),
    alpha = 0.45,
    inherit.aes = FALSE) +
  geom_errorbar(aes(ymin = ci_lo, ymax = ci_hi), width = 0, linewidth = 0.8) +
  geom_point(shape = 21, fill = "white", size = 1.5, stroke = 1.8) +
  scale_x_continuous(
    breaks   = 1:3,
    labels   = levels(fig_rq3b_corr_ci_data$need),
    limits   = c(0.6, 3.6),
    position = "top"
  ) +
  scale_color_manual(
    values = all_color_values, name = NULL, limits = all_levels,
    labels = c(LABEL_HU, LABEL_AI, "Δ<i>r</i>"),
    guide  = guide_legend(nrow = 2, byrow = TRUE, override.aes = list(
      shape     = c(21L,  21L,  22L),
      linetype  = c("solid", "solid", "blank"),
      linewidth = c(0.8,  0.8,  0),
      size      = c(1.5,  1.5,  3.5),
      fill      = c("white", "white", COL_DIFF),
      color     = c(rq3b_group_colors_ci[[LABEL_HU]],
                    rq3b_group_colors_ci[[LABEL_AI]],
                    "transparent"),
      alpha     = c(1, 1, 0.45)
    ))
  ) +
  scale_fill_manual(values = all_fill_values, name = NULL, limits = all_levels,
    guide = "none") +
  labs(x = NULL, y = "Pearson *r*") +
  theme_thesis() +
  theme(
    panel.grid.major.x   = element_blank(),
    panel.grid.minor.x   = element_blank(),
    axis.text.x.top      = element_text(colour = "black"),
    panel.grid.major.y   = element_line(linewidth = 0.3),
    panel.grid.minor.y   = element_line(linewidth = 0.3),
    axis.title.y         = ggtext::element_markdown(family = THESIS_FONT, margin = margin(r = 12)),
    legend.position      = c(0.23, 0.13),
    legend.direction     = "horizontal",
    legend.margin        = margin(t = 4),
    legend.background    = element_rect(fill = "white", color = NA),
    legend.key           = element_blank(),
    legend.text          = ggtext::element_markdown(family = THESIS_FONT)
  )
```

*Note.* Error bars = 95% bootstrap CIs (R = `r BOOTSTRAP_SAMPLES`). Shaded rectangles = Δr = r(AI) − r(Human). AI surrogates N = `r N_AI`; Humans N = `r N_HU`.

## 5.6 Spearman Robustness

```{r}
#| label: compute-rq3b-spearman
#| include: false
#| cache: true
#| cache.extra: !expr list(file.mtime(rds_paths$merged), POLY_BOOT_R_RQ2)

boot_fn_delta_rho <- function(data, indices) {
  d  <- data[indices, ]
  ai <- dplyr::filter(d, id_type == "AI")
  hu <- dplyr::filter(d, id_type == "HU")
  c(
    aut = cor(ai$llm_aut_mean, ai$q_aut_mean, use = "complete.obs", method = "spearman") -
          cor(hu$llm_aut_mean, hu$q_aut_mean, use = "complete.obs", method = "spearman"),
    com = cor(ai$llm_com_mean, ai$q_com_mean, use = "complete.obs", method = "spearman") -
          cor(hu$llm_com_mean, hu$q_com_mean, use = "complete.obs", method = "spearman"),
    rel = cor(ai$llm_rel_mean, ai$q_rel_mean, use = "complete.obs", method = "spearman") -
          cor(hu$llm_rel_mean, hu$q_rel_mean, use = "complete.obs", method = "spearman")
  )
}

set.seed(2026)
boot_spearman_rq3 <- boot::boot(
  data      = d_rq3_merged,
  statistic = boot_fn_delta_rho,
  R         = POLY_BOOT_R_RQ2,
  strata    = as.integer(factor(d_rq3_merged$id_type))
)
colnames(boot_spearman_rq3$t) <- names(boot_spearman_rq3$t0)

message("boot_spearman_rq3 complete.")
```

```{r}
#| label: tbl-rq3b-spearman
#| echo: false
#| tbl-cap: "RQ3b Spearman Robustness: Rho per Group and Delta-rho with Percentile CIs"

purrr::map_dfr(
  list(list(llm = "llm_aut_mean", q = "q_aut_mean", need = TABLE_LABEL_AUT, idx = 1L),
       list(llm = "llm_com_mean", q = "q_com_mean", need = TABLE_LABEL_COM, idx = 2L),
       list(llm = "llm_rel_mean", q = "q_rel_mean", need = TABLE_LABEL_REL, idx = 3L)),
  function(x) {
    ai_dat <- d_rq3_merged[d_rq3_merged$id_type == "AI", ]
    hu_dat <- d_rq3_merged[d_rq3_merged$id_type == "HU", ]
    rho_ai <- cor(ai_dat[[x$llm]], ai_dat[[x$q]], use = "complete.obs", method = "spearman")
    rho_hu <- cor(hu_dat[[x$llm]], hu_dat[[x$q]], use = "complete.obs", method = "spearman")
    d_rho_ci <- perc_ci(boot_spearman_rq3$t[, x$idx])
    tibble::tibble(
      Need        = x$need,
      rho_AI      = round(rho_ai, 2),
      rho_HU      = round(rho_hu, 2),
      delta_rho   = round(rho_ai - rho_hu, 2),
      delta_rho_ci = sprintf("[%.2f, %.2f]", d_rho_ci[1], d_rho_ci[2])
    )
  }
) |>
  flextable() |>
  flextable::set_header_labels(
    rho_AI = "ρ (AI)", rho_HU = "ρ (HU)",
    delta_rho = "Δ ρ", delta_rho_ci = "95% CI (Δ ρ)"
  ) |>
  style_apa7() |>
  add_apa_note(paste0(
    "Spearman rho on unrounded continuous scores. ",
    "Δ ρ bootstrap percentile CIs, R = ", POLY_BOOT_R_RQ2, ", stratified by group. ",
    "Compare with Pearson r in tbl-rq3b-results."
  ))
```

## 5.7 RQ3b Delta-r Forest Plot

```{r}
#| label: fig-rq3b-delta-r-forest
#| echo: false
#| fig-cap: "RQ3b: Bootstrap Estimates of Δ r (AI − Human) by Need Dimension. Points = observed Pearson r difference; horizontal segments = 95% bootstrap percentile CIs; dashed line = 0."
#| fig-width: 6
#| fig-height: 3
#| fig-dpi: 300

delta_r_forest_data <- tibble::tibble(
  Need     = factor(NEEDS, levels = rev(NEEDS)),
  estimate = as.numeric(boot_delta_r$t0),
  ci_lo    = apply(boot_delta_r$t, 2L, quantile, CI_Q_LO, na.rm = TRUE),
  ci_hi    = apply(boot_delta_r$t, 2L, quantile, CI_Q_HI, na.rm = TRUE)
)

ggplot(delta_r_forest_data, aes(x = estimate, y = Need)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  geom_segment(aes(x = ci_lo, xend = ci_hi, y = Need, yend = Need), linewidth = 0.8) +
  geom_point(size = 3, shape = 21, fill = "white", stroke = 1.8) +
  labs(x = "Δ r (AI − Human) [95% bootstrap CI]", y = NULL) +
  theme_thesis() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(linewidth = 0.2)
  )
```

---

# 6 Appendix: BCa Sensitivity Tables

```{r}
#| label: tbl-bca-delta-m-appendix
#| echo: false
#| tbl-cap: "Appendix: BCa vs Percentile CIs for Delta-M — All 3 Need Dimensions"

bca_delta_m_results |>
  dplyr::mutate(dplyr::across(where(is.numeric), \(x) round(x, 3))) |>
  flextable() |>
  style_apa7() |>
  add_apa_note("Divergence threshold = 0.02. Cells with max_diff > 0.02 also reported in check_bca_divergence_delta_m.")
```

```{r}
#| label: tbl-bca-delta-sd-appendix
#| echo: false
#| tbl-cap: "Appendix: BCa vs Percentile CIs for Delta-SD — All 3 Need Dimensions"

bca_delta_sd_results |>
  dplyr::mutate(dplyr::across(where(is.numeric), \(x) round(x, 3))) |>
  flextable() |>
  style_apa7() |>
  add_apa_note("Divergence threshold = 0.02. Cells with max_diff > 0.02 also reported in check_bca_divergence_delta_sd.")
```

```{r}
#| label: tbl-bca-delta-r-appendix
#| echo: false
#| tbl-cap: "Appendix: BCa vs Percentile CIs for Delta-r — All 3 Need Dimensions"

bca_delta_r_results |>
  dplyr::mutate(dplyr::across(where(is.numeric), \(x) round(x, 3))) |>
  flextable() |>
  style_apa7() |>
  add_apa_note("Divergence threshold = 0.02. Cells with max_diff > 0.02 also reported in check_bca_divergence_delta_r.")
```

---

# 7 Session Info

```{r}
#| label: session-info
#| include: false

sessionInfo()
```
