---
title: "Additional Analyses — H1c/H1d: Human Discriminant Validity and Structural Consistency"
subtitle: "Steiger's Test · Meta-r · Spearman Robustness · AI Surrogate Context"
author: "Caroline Müller"
date: "2026-05-17"
format:
  html:
    code-fold: true
    code-summary: "Show code"
    code-tools: true
    toc: true
    self-contained: true
    fig-cap-location: top
editor:
  markdown:
    wrap: 72
knitr:
  opts_chunk:
    dpi: 300
---

# 1 Overview

**Prerequisites:** Run the following files first:

- `10_RQ2_data.qmd` → `d_rq3_ai_conditions.rds`
- `20_RQ2_analyses.qmd` → `h1c_ai_results.rds`, `h1d_ai_results.rds`
- `30_RQ3_data.qmd` → `d_human_clean.rds`, `d_rq3_merged.rds`

This file covers H1c (discriminant validity) and H1d (structural
consistency) for the **human** participant sample, with AI surrogate
results loaded from pre-computed RDS files as context. Sections 3 and 4
follow a parallel structure:

1. **AI surrogate context** — loaded from RDS (all 9 conditions)
2. **Human analysis** — computed here (matched short-3 condition)
3. **Comparison** — human vs AI short-3 condition side by side
4. **Spearman robustness** — same tests with Spearman ρ as approximation

**Note on questionnaire items:** Human Q-Relatedness is scored from 5
items (`q_rel_1`–`q_rel_5`), while AI Q-Relatedness uses 4 items.
This difference is inherited from the data preparation step and is
explicitly noted wherever it affects interpretation.

**Note on human condition:** All human participants were assessed in the
matched short-3 condition (target words = 50, text count = 3). H1c and
H1d results for humans are therefore based on a single condition.

---

# 2 Setup

## 2.1 Environment

```{r}
#| label: setup-environment
#| include: false
#| warning: false

graphics.off()

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(purrr)
  library(ggplot2)
  library(flextable)
  library(officer)
  library(scales)
  library(systemfonts)
  library(ragg)
  library(cocor)
})

knitr::opts_chunk$set(dev = "ragg_png")
if (.Platform$OS.type == "windows") windowsFonts(Calibri = windowsFont("Calibri"))

message("Libraries loaded.")
```

## 2.2 Shared Constants

```{r}
#| label: setup-common
#| include: false

source("00_common.R")
```

## 2.3 Load Upstream Exports

```{r}
#| label: setup-imports
#| include: false

EXPORT_DIR <- "../10_data/80_analyses_exports"

rds_paths <- list(
  human_clean = file.path(EXPORT_DIR, "d_human_clean.rds"),
  merged      = file.path(EXPORT_DIR, "d_rq3_merged.rds"),
  h1c_ai      = file.path(EXPORT_DIR, "h1c_ai_results.rds"),
  h1d_ai      = file.path(EXPORT_DIR, "h1d_ai_results.rds"),
  ai_cond     = file.path(EXPORT_DIR, "d_rq3_ai_conditions.rds")
)

for (nm in names(rds_paths)) {
  if (!file.exists(rds_paths[[nm]])) {
    stop("Missing upstream export: ", rds_paths[[nm]],
         "\nRun the prerequisite QMD files listed in Section 1.")
  }
}

d_human_clean        <- readRDS(rds_paths$human_clean)
d_rq3_merged         <- readRDS(rds_paths$merged)
h1c_ai_results       <- readRDS(rds_paths$h1c_ai)
h1d_ai_results       <- readRDS(rds_paths$h1d_ai)
d_rq3_ai_conditions  <- readRDS(rds_paths$ai_cond)

N_HU         <- nrow(d_human_clean)
N_AI_MATCHED <- sum(d_rq3_merged$id_type == "AI")

message("Imports loaded: ", N_HU, " human rows; ",
        N_AI_MATCHED, " AI short-3 rows; ",
        nrow(h1c_ai_results), " AI H1c rows; ",
        nrow(h1d_ai_results), " AI H1d rows; ",
        nrow(d_rq3_ai_conditions), " AI condition rows.")
```

---

# 3 H1c — Discriminant Validity

H1c requires that each LLM score dimension correlates more strongly with
its matching questionnaire dimension (convergent, r_within) than with
non-matching dimensions (discriminant, r_cross). Steiger's (1980) test
for dependent overlapping correlations (`cocor::cocor.dep.groups.overlap`)
is used because all three correlation inputs share the same LLM score.
Six comparisons per sample: each of three LLM dimensions compared against
both non-matching Q dimensions. All tests two-tailed.

The Steiger helpers (`run_steiger_cond`, `run_steiger_cond_spearman`) are
defined in `00_common.R`. CI SE is back-computed from the cocor z-statistic
to ensure consistency between the p-value and the interval.

## 3.1 AI Surrogate Context

AI H1c results across all 9 conditions are loaded from
`h1c_ai_results.rds` (computed in `20_RQ2_analyses.qmd`). The MTMM plot
below shows the convergent vs discriminant r pattern for each condition
and LLM dimension.

```{r}
#| label: fig-h1c-ai-context-mtmm
#| echo: false
#| fig-cap: "H1c Context: MTMM Validity Coefficient Comparison — AI Surrogates, All 9 Conditions. Filled circle = convergent r (same need); open circles = discriminant r. Discriminant validity holds when the filled circle lies to the right of all open circles in a row."
#| fig-width: 10
#| fig-height: 6
#| fig-dpi: 300

h1c_ai_long <- dplyr::bind_rows(
  h1c_ai_results |>
    dplyr::group_by(condition, llm_dim) |>
    dplyr::slice(1) |>
    dplyr::ungroup() |>
    dplyr::transmute(
      condition, llm_dim,
      q_dim  = sub("^(Q_\\w+) vs.*", "\\1", comparison),
      r_type = "Convergent", r = r_within
    ),
  h1c_ai_results |>
    dplyr::transmute(
      condition, llm_dim,
      q_dim  = sub(".*vs (Q_\\w+)$", "\\1", comparison),
      r_type = "Discriminant", r = r_cross
    )
) |>
  dplyr::mutate(
    text_count   = as.integer(sub("^(\\d+) texts?.*", "\\1", condition)),
    target_words = as.integer(sub(".*(\\d+)w$",        "\\1", condition)),
    llm_f = factor(llm_dim,
                   levels = c("llm_aut_mean", "llm_com_mean", "llm_rel_mean"),
                   labels = c("LLM Autonomy", "LLM Competence", "LLM Relatedness")),
    q_f   = factor(q_dim,
                   levels = c("Q_Aut", "Q_Com", "Q_Rel"),
                   labels = c("Q Aut", "Q Com", "Q Rel")),
    cond_f = factor(
      paste0(text_count, "t × ", target_words, "w"),
      levels = paste0(rep(c(1L, 3L, 5L), each = 3L), "t × ",
                      rep(c(50L, 150L, 300L), 3L), "w")
    )
  )

ggplot(h1c_ai_long,
       aes(x = r, y = cond_f, group = interaction(r_type, q_f),
           shape = r_type, color = q_f, size = r_type)) +
  geom_vline(xintercept = 0, color = "grey70", linewidth = 0.3) +
  geom_point(position = position_dodge(width = 0.6), alpha = 0.9) +
  scale_shape_manual(values = c("Convergent" = 19L, "Discriminant" = 1L), name = NULL) +
  scale_size_manual(values = c("Convergent" = 3.5, "Discriminant" = 2.5), guide = "none") +
  scale_color_manual(
    values = c("Q Aut" = COL_AI_SHADES[1], "Q Com" = COL_AI_SHADES[2],
               "Q Rel" = COL_AI_SHADES[4]),
    name = "Q dimension"
  ) +
  facet_wrap(~llm_f, ncol = 3L) +
  labs(x = "Correlation (r)", y = NULL) +
  theme_thesis()
```

## 3.2 Human H1c Analysis

Six Steiger tests on `d_human_clean` (N = `r N_HU`). Human condition:
short-3 (target words = 50, text count = 3). **Note:** Q-Relatedness
for humans uses 5 items (`q_rel_1`–`q_rel_5`); AI uses 4. This applies
to both `q_rel_mean` (the composite mean used in all correlations below)
and to the inter-need correlations involving relatedness.

```{r}
#| label: compute-h1c-human
#| include: false

HU_COND_LABEL <- "Human (short-3)"

tbl_h1c_human <- dplyr::bind_rows(
  run_steiger_cond(d_human_clean, "llm_aut_mean", "q_aut_mean", "q_com_mean",
                   "Q_Aut", "Q_Com", HU_COND_LABEL),
  run_steiger_cond(d_human_clean, "llm_aut_mean", "q_aut_mean", "q_rel_mean",
                   "Q_Aut", "Q_Rel", HU_COND_LABEL),
  run_steiger_cond(d_human_clean, "llm_com_mean", "q_com_mean", "q_aut_mean",
                   "Q_Com", "Q_Aut", HU_COND_LABEL),
  run_steiger_cond(d_human_clean, "llm_com_mean", "q_com_mean", "q_rel_mean",
                   "Q_Com", "Q_Rel", HU_COND_LABEL),
  run_steiger_cond(d_human_clean, "llm_rel_mean", "q_rel_mean", "q_aut_mean",
                   "Q_Rel", "Q_Aut", HU_COND_LABEL),
  run_steiger_cond(d_human_clean, "llm_rel_mean", "q_rel_mean", "q_com_mean",
                   "Q_Rel", "Q_Com", HU_COND_LABEL)
)

message("H1c human: ", nrow(tbl_h1c_human), " Steiger tests computed.")
```

```{r}
#| label: tbl-h1c-human
#| echo: false
#| tbl-cap: "H1c: Discriminant Validity — Steiger's Test for Human Participants (Short-3 Condition, N = 6 Tests)"

tbl_h1c_human |>
  dplyr::select(-condition) |>
  flextable() |>
  flextable::set_header_labels(
    llm_dim    = "LLM dim",
    comparison = "Comparison",
    r_within   = "r_within",
    r_cross    = "r_cross",
    diff       = "Diff",
    ci_lo      = "CI lo",
    ci_hi      = "CI hi",
    steiger_z  = "z",
    p_value    = "p",
    n          = "N"
  ) |>
  style_apa7() |>
  add_apa_note(paste0(
    "Steiger (1980) test for dependent overlapping correlations ",
    "(cocor::cocor.dep.groups.overlap). Two-tailed. N = ", N_HU, ". ",
    "r_within = LLM-matching Q correlation (convergent); ",
    "r_cross = LLM-nonmatching Q correlation (discriminant). ",
    "CI = 95% Fisher-z CI on r_within − r_cross. ",
    "Q-Relatedness: 5 items for humans (vs 4 for AI surrogates)."
  ))
```

```{r}
#| label: fig-h1c-human-mtmm
#| echo: false
#| fig-cap: "H1c: MTMM Validity Coefficient Comparison — Human Participants (Short-3 Condition). Filled circle = convergent r; open circles = discriminant r. Discriminant validity holds when the filled circle lies to the right of all open circles in each panel."
#| fig-width: 8
#| fig-height: 3.5
#| fig-dpi: 300

h1c_human_long <- dplyr::bind_rows(
  tbl_h1c_human |>
    dplyr::group_by(llm_dim) |>
    dplyr::slice(1) |>
    dplyr::ungroup() |>
    dplyr::transmute(
      llm_dim,
      q_dim  = sub("^(Q_\\w+) vs.*", "\\1", comparison),
      r_type = "Convergent", r = r_within
    ),
  tbl_h1c_human |>
    dplyr::transmute(
      llm_dim,
      q_dim  = sub(".*vs (Q_\\w+)$", "\\1", comparison),
      r_type = "Discriminant", r = r_cross
    )
) |>
  dplyr::mutate(
    llm_f = factor(llm_dim,
                   levels = c("llm_aut_mean", "llm_com_mean", "llm_rel_mean"),
                   labels = c("LLM Autonomy", "LLM Competence", "LLM Relatedness")),
    q_f   = factor(q_dim,
                   levels = c("Q_Aut", "Q_Com", "Q_Rel"),
                   labels = c("Q Aut", "Q Com", "Q Rel"))
  )

ggplot(h1c_human_long,
       aes(x = r, y = q_f, shape = r_type, color = q_f, size = r_type)) +
  geom_vline(xintercept = 0, color = "grey70", linewidth = 0.3) +
  geom_point(alpha = 0.9) +
  scale_shape_manual(values = c("Convergent" = 19L, "Discriminant" = 1L), name = NULL) +
  scale_size_manual(values = c("Convergent" = 3.5, "Discriminant" = 2.5), guide = "none") +
  scale_color_manual(
    values = c("Q Aut" = COL_AI_SHADES[1], "Q Com" = COL_AI_SHADES[2],
               "Q Rel" = COL_AI_SHADES[4]),
    name = "Q dimension"
  ) +
  facet_wrap(~llm_f, ncol = 3L) +
  labs(x = "Correlation (r)", y = "Q dimension") +
  theme_thesis()
```

## 3.3 Comparison: Human vs AI Short-3

```{r}
#| label: tbl-h1c-compare
#| echo: false
#| tbl-cap: "H1c Comparison: Human vs AI Surrogates (Matched Short-3 Condition, target words = 50, text count = 3)"

# AI short-3 rows: filter by condition matching "3 texts × 50w"
AI_SHORT3_LABEL <- paste0("3 texts × 50w")  # "3 texts × 50w" — must match label built in 20_RQ2_analyses.qmd
h1c_ai_short3 <- h1c_ai_results |>
  dplyr::filter(condition == AI_SHORT3_LABEL)

tbl_h1c_compare <- dplyr::bind_rows(
  h1c_ai_short3   |> dplyr::mutate(Group = "AI (short-3)"),
  tbl_h1c_human   |> dplyr::mutate(Group = "Human (short-3)")
) |>
  dplyr::select(Group, llm_dim, comparison, r_within, r_cross,
                diff, ci_lo, ci_hi, steiger_z, p_value, n) |>
  dplyr::arrange(comparison, Group)

tbl_h1c_compare |>
  flextable() |>
  flextable::set_header_labels(
    Group      = "Group",
    llm_dim    = "LLM dim",
    comparison = "Comparison",
    r_within   = "r_within",
    r_cross    = "r_cross",
    diff       = "Diff",
    ci_lo      = "CI lo",
    ci_hi      = "CI hi",
    steiger_z  = "z",
    p_value    = "p",
    n          = "N"
  ) |>
  style_apa7() |>
  add_apa_note(paste0(
    "AI short-3 results from h1c_ai_results.rds (computed in 20_RQ2_analyses.qmd). ",
    "Human short-3 results computed above (Section 3.2). ",
    "Q-Relatedness: AI = 4 items; Human = 5 items."
  ))
```

## 3.4 Spearman Robustness

Steiger's (1980) test is designed for Pearson r; applying it to Spearman
ρ is an approximation. All other test parameters are identical to
Section 3.2.

```{r}
#| label: compute-h1c-human-spearman
#| include: false

tbl_h1c_human_spearman <- dplyr::bind_rows(
  run_steiger_cond_spearman(d_human_clean, "llm_aut_mean", "q_aut_mean", "q_com_mean",
                            "Q_Aut", "Q_Com", HU_COND_LABEL),
  run_steiger_cond_spearman(d_human_clean, "llm_aut_mean", "q_aut_mean", "q_rel_mean",
                            "Q_Aut", "Q_Rel", HU_COND_LABEL),
  run_steiger_cond_spearman(d_human_clean, "llm_com_mean", "q_com_mean", "q_aut_mean",
                            "Q_Com", "Q_Aut", HU_COND_LABEL),
  run_steiger_cond_spearman(d_human_clean, "llm_com_mean", "q_com_mean", "q_rel_mean",
                            "Q_Com", "Q_Rel", HU_COND_LABEL),
  run_steiger_cond_spearman(d_human_clean, "llm_rel_mean", "q_rel_mean", "q_aut_mean",
                            "Q_Rel", "Q_Aut", HU_COND_LABEL),
  run_steiger_cond_spearman(d_human_clean, "llm_rel_mean", "q_rel_mean", "q_com_mean",
                            "Q_Rel", "Q_Com", HU_COND_LABEL)
)

message("H1c human Spearman: ", nrow(tbl_h1c_human_spearman), " tests computed.")
```

```{r}
#| label: tbl-h1c-human-spearman
#| echo: false
#| tbl-cap: "H1c Spearman Robustness: Steiger's Test with Spearman rho — Human Participants (Short-3 Condition; Approximation)"

tbl_h1c_human_spearman |>
  dplyr::select(-condition) |>
  flextable() |>
  flextable::set_header_labels(
    llm_dim    = "LLM dim",
    comparison = "Comparison",
    rho_within = "rho_within",
    rho_cross  = "rho_cross",
    diff       = "Diff",
    ci_lo      = "CI lo",
    ci_hi      = "CI hi",
    steiger_z  = "z",
    p_value    = "p",
    n          = "N"
  ) |>
  style_apa7() |>
  add_apa_note(paste0(
    "Steiger (1980) applied to Spearman rho — approximation; ",
    "compare with Pearson results in tbl-h1c-human. N = ", N_HU, "."
  ))
```

---

# 4 H1d — Structural Consistency

H1d asks whether the inter-need correlation pattern in LLM scores
mirrors the pattern in questionnaire scores. Three inter-need Pearson r
values are computed for each source (Q and LLM): Aut–Com, Aut–Rel,
Com–Rel. The meta-correlation (meta-r) is the Pearson r between these
two 3-element vectors. N = 3 data points — no inferential test is
applied; results are descriptive only.

**Note:** Q inter-need correlations for AI surrogates are constant
across conditions (Q scores come from one questionnaire administration
per persona, independent of scoring condition). For humans there is only
one condition (short-3), so meta-r is a single value.

## 4.1 AI Surrogate Context

AI meta-r values per condition (from `h1d_ai_results.rds`) are shown
below to provide context before the human result. Condition-to-condition
variation reflects shifts in the LLM's inter-need structure across text
length and count.

```{r}
#| label: tbl-h1d-ai-summary
#| echo: false
#| tbl-cap: "H1d Context: AI Surrogate Meta-r by Condition (N = 9 Conditions, Descriptive Only)"

h1d_ai_results |>
  dplyr::distinct(condition, meta_r) |>
  dplyr::mutate(
    text_count   = as.integer(sub("^(\\d+) texts?.*", "\\1", condition)),
    target_words = as.integer(sub(".*(\\d+)w$", "\\1", condition))
  ) |>
  dplyr::arrange(text_count, target_words) |>
  dplyr::select(Condition = condition, `Meta-r` = meta_r) |>
  flextable() |>
  style_apa7() |>
  add_apa_note(paste0(
    "meta_r = Pearson r between three Q inter-need correlations and three LLM ",
    "inter-need correlations per condition. N = 3 pairs; descriptive only. ",
    "Loaded from h1d_ai_results.rds (computed in 20_RQ2_analyses.qmd)."
  ))
```

```{r}
#| label: fig-h1d-ai-context-forest
#| echo: false
#| fig-cap: "H1d Context: Inter-Need Correlations by Instrument, Source, and Condition (Pearson r, 95% CI). Each row is one source: human participants (top, separated by dashed line) or one of the nine AI surrogate conditions. Within each row, two points show the questionnaire inter-need correlation (Q, amber) and the LLM inter-need correlation (LLM, teal). Error bars = 95% CI via Fisher z-transform. AI questionnaire correlations are constant across conditions (same personas, same Q scores). Descriptive only."
#| fig-width: 10
#| fig-height: 6
#| fig-dpi: 300

# ── helper: Pearson r with 95% CI (Fisher z) ─────────────────────────────────
.r_ci <- function(x, y, conf = 0.95) {
  cc   <- complete.cases(x, y)
  n    <- sum(cc)
  r    <- cor(x[cc], y[cc])
  z    <- atanh(r)
  se   <- 1 / sqrt(n - 3)
  crit <- qnorm((1 + conf) / 2)
  data.frame(r = r, ci_lo = tanh(z - crit * se), ci_hi = tanh(z + crit * se), n = n)
}

.pairs <- list(
  list(suf1 = "aut", suf2 = "com", label = "Autonomy–Competence"),
  list(suf1 = "aut", suf2 = "rel", label = "Autonomy–Relatedness"),
  list(suf1 = "com", suf2 = "rel", label = "Competence–Relatedness")
)

# ── AI conditions ─────────────────────────────────────────────────────────────
.ai_grid <- d_rq3_ai_conditions |>
  dplyr::distinct(target_words, text_count) |>
  dplyr::arrange(text_count, target_words)

.ai_rows <- purrr::map_dfr(seq_len(nrow(.ai_grid)), function(i) {
  lbl <- paste0(.ai_grid$text_count[i], "t × ", .ai_grid$target_words[i], "w")
  sub <- dplyr::filter(d_rq3_ai_conditions,
                       target_words == .ai_grid$target_words[i],
                       text_count   == .ai_grid$text_count[i])
  purrr::map_dfr(.pairs, function(p) {
    dplyr::bind_rows(
      .r_ci(sub[[paste0("q_",   p$suf1, "_mean")]],
            sub[[paste0("q_",   p$suf2, "_mean")]]) |>
        dplyr::mutate(instrument = "Questionnaire"),
      .r_ci(sub[[paste0("llm_", p$suf1, "_mean")]],
            sub[[paste0("llm_", p$suf2, "_mean")]]) |>
        dplyr::mutate(instrument = "LLM Score")
    ) |> dplyr::mutate(source = lbl, source_type = "AI Surrogate", pair = p$label)
  })
})

# ── Human ─────────────────────────────────────────────────────────────────────
.hu_rows <- purrr::map_dfr(.pairs, function(p) {
  dplyr::bind_rows(
    .r_ci(d_human_clean[[paste0("q_",   p$suf1, "_mean")]],
          d_human_clean[[paste0("q_",   p$suf2, "_mean")]]) |>
      dplyr::mutate(instrument = "Questionnaire"),
    .r_ci(d_human_clean[[paste0("llm_", p$suf1, "_mean")]],
          d_human_clean[[paste0("llm_", p$suf2, "_mean")]]) |>
      dplyr::mutate(instrument = "LLM Score")
  ) |> dplyr::mutate(source = "Human (short-3)", source_type = "Human", pair = p$label)
})

# ── factor ordering: AI bottom-to-top, Human at top ──────────────────────────
.ai_levels <- paste0(rep(c(1L, 3L, 5L), each = 3L), "t × ",
                     rep(c(50L, 150L, 300L), 3L), "w")

.forest_data <- dplyr::bind_rows(.ai_rows, .hu_rows) |>
  dplyr::mutate(
    source     = factor(source, levels = c(.ai_levels, "Human (short-3)")),
    pair_f     = factor(pair, levels = c("Autonomy–Competence",
                                         "Autonomy–Relatedness",
                                         "Competence–Relatedness")),
    instrument = factor(instrument, levels = c("Questionnaire", "LLM Score"))
  )

# ── plot ──────────────────────────────────────────────────────────────────────
COL_INSTR  <- c("Questionnaire" = "#C47722", "LLM Score" = "#1F5A71")
SHAPE_INSTR <- c("Questionnaire" = 15L,      "LLM Score" = 16L)  # square / circle

ggplot(.forest_data,
       aes(x = r, y = source, xmin = ci_lo, xmax = ci_hi,
           colour = instrument, shape = instrument)) +
  geom_vline(xintercept = 0, colour = "grey70", linewidth = 0.3) +
  geom_hline(yintercept = 9.5, colour = "grey50", linewidth = 0.35,
             linetype = "dashed") +
  geom_errorbarh(aes(group = instrument),
                 height = 0.25, linewidth = 0.5, alpha = 0.8,
                 position = position_dodge(width = 0.55)) +
  geom_point(size = 2.5,
             position = position_dodge(width = 0.55)) +
  facet_wrap(~pair_f, ncol = 3L) +
  scale_colour_manual(values = COL_INSTR,  name = "Instrument") +
  scale_shape_manual( values = SHAPE_INSTR, name = "Instrument") +
  scale_x_continuous(breaks = seq(0, 0.8, 0.2)) +
  labs(x = "Pearson r (95% CI)", y = NULL) +
  theme_thesis() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(linewidth = 0.2, colour = "grey88"),
    legend.position    = "top",
    axis.text.y        = element_text(size = 9)
  )
```

```{r}
#| label: compute-meta-r-ai
#| include: false

meta_r_ai <- h1d_ai_results |>
  dplyr::distinct(condition, meta_r) |>
  dplyr::mutate(
    text_count   = as.integer(sub("^(\\d+) texts?.*", "\\1", condition)),
    target_words = as.integer(sub(".*(\\d+)w$", "\\1", condition))
  )
```

## 4.2 Human H1d Analysis

```{r}
#| label: compute-h1d-human
#| include: false

# Inter-need Pearson r for Q composites (5-item relatedness for humans)
q_vec_hu <- c(
  cor(d_human_clean$q_aut_mean, d_human_clean$q_com_mean, use = "complete.obs"),
  cor(d_human_clean$q_aut_mean, d_human_clean$q_rel_mean, use = "complete.obs"),
  cor(d_human_clean$q_com_mean, d_human_clean$q_rel_mean, use = "complete.obs")
)

# Inter-need Pearson r for LLM composites
llm_vec_hu <- c(
  cor(d_human_clean$llm_aut_mean, d_human_clean$llm_com_mean, use = "complete.obs"),
  cor(d_human_clean$llm_aut_mean, d_human_clean$llm_rel_mean, use = "complete.obs"),
  cor(d_human_clean$llm_com_mean, d_human_clean$llm_rel_mean, use = "complete.obs")
)

meta_r_human <- round(cor(q_vec_hu, llm_vec_hu), 2)

tbl_h1d_human <- data.frame(
  pair      = c("Aut-Com", "Aut-Rel", "Com-Rel"),
  r_Q       = round(q_vec_hu,   3),
  r_LLM     = round(llm_vec_hu, 2),
  meta_r    = c(meta_r_human, NA_real_, NA_real_)
)

message("H1d human: meta_r = ", meta_r_human)
```

```{r}
#| label: tbl-h1d-human
#| echo: false
#| tbl-cap: "H1d: Structural Consistency — Inter-Need Correlations and Meta-r for Human Participants (Short-3 Condition; N = 3 Pairs, Descriptive Only)"

tbl_h1d_human |>
  flextable() |>
  flextable::colformat_num(j = "meta_r", na_str = "") |>
  flextable::set_header_labels(
    pair   = "Need pair",
    r_Q    = "r (Q)",
    r_LLM  = "r (LLM)",
    meta_r = "Meta-r"
  ) |>
  style_apa7() |>
  add_apa_note(paste0(
    "r (Q) = Pearson r between Q composite means (human sample). ",
    "Q-Relatedness uses 5 items (vs 4 for AI). ",
    "meta_r = Pearson r between the three Q r-values and three LLM r-values ",
    "(condition-level statistic, shown once). N = 3 pairs; descriptive only. ",
    "Sample: N = ", N_HU, " human participants."
  ))
```

## 4.3 Comparison: Human vs AI Short-3

LLM inter-need correlations for the matched AI short-3 condition (3
texts × 50w) and the human short-3 result are shown side by side per
need pair. Dashed grey line = AI Q inter-need r; dotted red line = Human
Q inter-need r (5-item relatedness composite).

```{r}
#| label: fig-h1d-compare-forest
#| echo: false
#| fig-cap: "H1d Comparison: LLM Inter-Need Correlations — AI Surrogate Short-3 vs Human Short-3. Each point = one group. AI surrogate: filled circle (blue); Human: filled diamond (red). Dashed grey line = AI Q inter-need r; dotted red line = Human Q inter-need r (5-item relatedness). Descriptive only; N = 3 pairs per source."
#| fig-width: 9
#| fig-height: 3.5
#| fig-dpi: 300

fig_h1d_cmp <- dplyr::bind_rows(
  h1d_ai_results |>
    dplyr::filter(condition == "3 texts × 50w") |>
    dplyr::transmute(
      pair, r_LLM, r_Q,
      cond_label = "AI (short-3)",
      source     = "AI"
    ),
  tbl_h1d_human |>
    dplyr::transmute(
      pair, r_LLM, r_Q,
      cond_label = "Human (short-3)",
      source     = "Human"
    )
) |>
  dplyr::mutate(
    pair_f   = factor(pair,
                      levels = c("Aut-Com", "Aut-Rel", "Com-Rel"),
                      labels = c("Autonomy–Competence",
                                 "Autonomy–Relatedness",
                                 "Competence–Relatedness")),
    cond_f   = factor(cond_label,
                      levels = c("AI (short-3)", "Human (short-3)")),
    source_f = factor(source, levels = c("AI", "Human"))
  )

q_refs_cmp <- fig_h1d_cmp |>
  dplyr::distinct(pair_f, source, r_Q)

col_cmp <- c("AI" = COL_AI_SINGLE, "Human" = COL_HU_SINGLE)

ggplot(fig_h1d_cmp, aes(x = r_LLM, y = cond_f, color = source_f)) +
  geom_vline(xintercept = 0, color = "grey80", linewidth = 0.4) +
  geom_vline(data = dplyr::filter(q_refs_cmp, source == "AI"),
             aes(xintercept = r_Q),
             linetype = "dashed", color = "grey40", linewidth = 0.6,
             inherit.aes = FALSE) +
  geom_vline(data = dplyr::filter(q_refs_cmp, source == "Human"),
             aes(xintercept = r_Q),
             linetype = "dotted", color = COL_HU_SINGLE, linewidth = 0.6,
             inherit.aes = FALSE) +
  geom_point(aes(size = source_f, shape = source_f)) +
  facet_wrap(~pair_f, ncol = 3L, scales = "free_x") +
  scale_color_manual(values = col_cmp, name = NULL) +
  scale_shape_manual(values = c("AI" = 16L, "Human" = 18L), name = NULL) +
  scale_size_manual(values  = c("AI" = 3.0, "Human" = 3.5), guide = "none") +
  labs(x = "Inter-need correlation (r)", y = NULL) +
  theme_thesis() +
  theme(
    panel.grid.major.y = element_line(linewidth = 0.2, color = "grey88"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    legend.position    = "top"
  )
```

## 4.4 Spearman Robustness

Spearman ρ is substituted for Pearson r in all six inter-need
correlations. The meta-r is still Pearson (ranking 3 ρ values is not
meaningful), so this tests whether the structural pattern result is
robust to rank-based vs linear associations. **Note:** With N = 3
pairs, a single pair's rank-order shift can change meta-r by ±0.5 or
more; the comparison between Pearson and Spearman meta-r values has
limited diagnostic power and should be interpreted descriptively.

```{r}
#| label: compute-h1d-human-spearman
#| include: false

q_vec_hu_sp <- c(
  cor(d_human_clean$q_aut_mean, d_human_clean$q_com_mean,
      use = "complete.obs", method = "spearman"),
  cor(d_human_clean$q_aut_mean, d_human_clean$q_rel_mean,
      use = "complete.obs", method = "spearman"),
  cor(d_human_clean$q_com_mean, d_human_clean$q_rel_mean,
      use = "complete.obs", method = "spearman")
)

llm_vec_hu_sp <- c(
  cor(d_human_clean$llm_aut_mean, d_human_clean$llm_com_mean,
      use = "complete.obs", method = "spearman"),
  cor(d_human_clean$llm_aut_mean, d_human_clean$llm_rel_mean,
      use = "complete.obs", method = "spearman"),
  cor(d_human_clean$llm_com_mean, d_human_clean$llm_rel_mean,
      use = "complete.obs", method = "spearman")
)

meta_r_human_sp <- round(cor(q_vec_hu_sp, llm_vec_hu_sp), 2)

tbl_h1d_human_spearman <- data.frame(
  pair      = c("Aut-Com", "Aut-Rel", "Com-Rel"),
  rho_Q     = round(q_vec_hu_sp,   3),
  rho_LLM   = round(llm_vec_hu_sp, 2),
  meta_r    = c(meta_r_human_sp, NA_real_, NA_real_)
)

message("H1d human Spearman: meta_r = ", meta_r_human_sp)
```

```{r}
#| label: tbl-h1d-human-spearman
#| echo: false
#| tbl-cap: "H1d Spearman Robustness: Inter-Need Rho and Meta-r for Human Participants (Short-3 Condition; N = 3 Pairs, Descriptive Only)"

tbl_h1d_human_spearman |>
  flextable() |>
  flextable::colformat_num(j = "meta_r", na_str = "") |>
  flextable::set_header_labels(
    pair    = "Need pair",
    rho_Q   = "rho (Q)",
    rho_LLM = "rho (LLM)",
    meta_r  = "Meta-r"
  ) |>
  style_apa7() |>
  add_apa_note(paste0(
    "Spearman rho for all six inter-need correlations; meta-r = Pearson r of the ",
    "resulting rho vectors. Compare with Pearson results in tbl-h1d-human. ",
    "N = ", N_HU, " human participants."
  ))
```

---

# 5 Supplementary Cross-Group Comparisons

`d_rq3_merged` (AI short-3 + human, matched condition) is used for
cross-group descriptive comparisons. The table below provides a quick
summary of LLM score means by group and need dimension.

```{r}
#| label: tbl-merged-descriptives
#| echo: false
#| tbl-cap: "Supplementary: LLM Score Descriptives by Group and Need Dimension (Matched Short-3 Condition)"

d_rq3_merged |>
  tidyr::pivot_longer(
    cols      = c(llm_aut_mean, llm_com_mean, llm_rel_mean),
    names_to  = "variable",
    values_to = "score"
  ) |>
  dplyr::mutate(
    Need  = factor(dplyr::case_when(
      grepl("aut", variable) ~ TABLE_LABEL_AUT,
      grepl("com", variable) ~ TABLE_LABEL_COM,
      grepl("rel", variable) ~ TABLE_LABEL_REL
    ), levels = NEEDS),
    Group = c(AI = LABEL_AI, HU = LABEL_HU)[as.character(id_type)]
  ) |>
  dplyr::group_by(Group, Need) |>
  dplyr::summarise(
    N  = sum(!is.na(score)),
    M  = round(mean(score, na.rm = TRUE), 2),
    SD = round(sd(score,   na.rm = TRUE), 2),
    .groups = "drop"
  ) |>
  flextable() |>
  style_apa7() |>
  add_apa_note(paste0(
    "LLM composite means (3 repeats averaged). ",
    "AI: matched short-3 condition (target words = 50, text count = 3). ",
    "Human: short-3 condition."
  ))
```

---

# 6 Session Info

```{r}
#| label: session-info
#| include: false

sessionInfo()
```
